---
title: "14_soleboldo_tabib_combined3D"
author: "BH"
date: "2024-03-21"
output:
 pdf_document:
    toc: yes
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load_packages}
library("dplyr")
library("Seurat")
library("patchwork")
library("readxl")
library("tidyseurat")
library("ggplot2")
library("tidyverse")
library("RCurl")
library("cowplot")
library("sctransform")
library("scPred")
library("Seurat")
library("magrittr")
library("caret")
library("dplyr")
library("janitor")
library("factoextra")
library("reshape2")

options(future.globals.maxSize = 12000 * 1024^2)

```
#Integrate Tabib and Sole seurat objects which contain 3D EP/DS senpred predictions 
```{r}
custom_theme <- readRDS(file = "custom_theme.rds")
querytabib3D <- readRDS(file = "querytabib3D.rds")
querysole3D <- readRDS(file = "querysole3D.rds")


list <- list(querytabib3D, querysole3D)

applySCTransformToList <- function(seuratList) {
  transformedList <- list()
  
  for (i in seq_along(seuratList)) {
    seuratObj <- seuratList[[i]]
    
    # Check if the object has already been SCTransformed
    if (!"SCT" %in% colnames(seuratObj)) {
      # Apply SCTransform if not already done
      seuratObj <- SCTransform(seuratObj)
    } else {
      message(paste("SCTransform already applied to Seurat object in list at index", i))
    }
    
    transformedList[[i]] <- seuratObj
  }
  
  return(transformedList)
}

# Apply SCTransform to each Seurat object in the list
transformed_seurat_list <- applySCTransformToList(list)

# Check the transformed objects
head(transformed_seurat_list[[1]]$RNA, n = 5)  # Replace "RNA" with your assay name
head(transformed_seurat_list[[2]]$RNA, n = 5)  # Replace "RNA" with your assay name

# Select integration features
integ_features <- SelectIntegrationFeatures(object.list = transformed_seurat_list, 
                                            nfeatures = 2000)

# Prepare for integration
list <- PrepSCTIntegration(object.list = transformed_seurat_list, 
                            anchor.features = integ_features)

# Find integration anchors
integ_anchors <- FindIntegrationAnchors(object.list = list, 
                                        normalization.method = "LogNormalize", 
                                        anchor.features = integ_features)

# Integrate data
seurat_integrated_3D <- IntegrateData(anchorset = integ_anchors, 
                                   normalization.method = "LogNormalize")


```

#Select highly variable features
```{r}
#Selecting highly variable features - high cell to cell variation
seurat_integrated_3D <- FindVariableFeatures(seurat_integrated_3D, selection.method ="vst", nfeatures=2000)
#Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(seurat_integrated_3D), 10)
top10
plot3 <- VariableFeaturePlot(seurat_integrated_3D)
#Plot variable features with labels
plot4 <- LabelPoints(plot=plot3, points = top10, repel=TRUE)
plot4

ggsave(
  "tabibsole_integrated_3D_variablefeatures.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/TabibSole_integrated"
  )
```
#Scale and perform PCA
```{r}
all.genes <- rownames(seurat_integrated_3D)
seurat_integrated_3D <- ScaleData(seurat_integrated_3D, features = all.genes)

#Performing a PCA
seurat_integrated_3D <- RunPCA(seurat_integrated_3D, features=VariableFeatures(object = seurat_integrated_3D))

#Plotting the PCA
PCA_no_label <- DimPlot(seurat_integrated_3D, reduction = "pca", seed = 123, combine=TRUE) + NoLegend()
PCA_Sample_label <- DimPlot(seurat_integrated_3D, reduction = "pca", seed = 123, group.by = "scpred_prediction")

PCA_no_label
PCA_Sample_label
```

```{r}
DimHeatmap(seurat_integrated_3D, dims=1, cells=500, balanced=TRUE)
DimHeatmap(seurat_integrated_3D, dims=1:15, cells=500, balanced=TRUE)
ElbowPlot(seurat_integrated_3D) 

#Suggests top 10 PCs explain most of the variance. Test this with jackstraw plot
#hdfseurat_jackstraw <- JackStraw(hdfseurat_filt_2D, num.replicate = 100)
#hdfseurat_jackstraw <- ScoreJackStraw(hdfseurat_jackstraw, dims = 1:20)
#JackStrawPlot(hdfseurat_jackstraw, dims = 1:15)
```

## Clustering the cells using KNN
```{r}
#Use K-nearest neighbours based on PCA distance in FindNeighbours() function based on previously definited dimensions (10) from elbow plot
seurat_integrated_3D <- FindNeighbors(seurat_integrated_3D, dims =1:10)

#group cells using findclusters function
seurat_integrated_3D <- FindClusters(seurat_integrated_3D, resolution = 0.4)

#Look at cluster ID's of first 5 cells
head(Idents(seurat_integrated_3D), 5)
```

## Plot UMAP
```{r}
#Running a UMAP
seurat_integrated_3D <- RunUMAP(seurat_integrated_3D, dims=1:10)

#Plotting UMAP with cluster IDs
DimPlot(seurat_integrated_3D, reduction = "umap", seed = 123) +
  custom_theme(base_size=20)

ggsave(
  "UMAP_tabibsoleintegrated3D.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/TabibSole_Integrated"
  )

#Colouring above UMAP plot with Sample type
DimPlot(seurat_integrated_3D, reduction = "umap",group.by = "scpred_prediction", seed=123) +
  custom_theme(base_size=20)

ggsave(
  "UMAP_tabibsoleintegrated3D_samplecolour.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/TabibSole_Integrated"
  )

#Colouring above UMAP plot with Sample type
DimPlot(seurat_integrated_3D, reduction = "umap",group.by = "orig.ident", seed=123) 

DimPlot(seurat_integrated_3D, reduction = "umap",group.by = "age", seed=123) 

DimPlot(seurat_integrated_3D, reduction = "umap",group.by = "age_group", seed=123) 

```
# Generating a correlation matrix of a panel of selected senescence genes - DS predicted cells
```{r}
# Subset the Seurat object based on scpred_prediction
filtered_seurat <- subset(seurat_integrated_3D, subset = scpred_prediction == "DS3D")

# Extract normalised gene counts for filtered cells
gene_counts <- filtered_seurat@assays$RNA@data

# Define gene list 
gene_list <- c("CDKN1A", "CDKN2A", "IGFBP3", "SERPINE1", "IGF1", "CXCL12", "IL6", "IL8", "MMP3", "TP53", "IL1B")

# Check if gene names in gene_list are present in row names (genes)
missing_genes <- gene_list[!gene_list %in% rownames(gene_counts)]
if (length(missing_genes) > 0) {
  stop(paste("Genes missing in gene counts:", paste(missing_genes, collapse = ", ")))
}

# Filter gene counts based on the gene list
gene_counts <- gene_counts[gene_list, ]
gene_counts <- as.matrix(gene_counts)
# Transpose gene counts matrix
gene_counts_transposed <- t(gene_counts)
gene_counts_transposed <- as.matrix(gene_counts_transposed)

# Calculate correlation matrix
correlation_matrix <- cor(gene_counts_transposed)

# Print correlation matrix
print(correlation_matrix)

dendrogram_row <- as.dendrogram(hclust(dist(correlation_matrix), method = "ward.D2"))
dendrogram_col <- as.dendrogram(hclust(dist(t(correlation_matrix)), method = "ward.D2"))

# Define breaks and color palette
breaks <- unique(c(seq(-1,-0.0001,length=50),seq(-0.00001,0.01,length=50), seq(0.011,1,length=50)))
my_palette <- colorRampPalette(c("blue","white", "white", "red"))(length(breaks)-1)

library("gplots")
png(filename = "../Output/TabibSole_Integrated/Markercorrelation_DS_tabibsole.png", width = 800, height = 600)  # Set appropriate width and height


heatmap <- heatmap.2(correlation_matrix, 
                     col = my_palette, 
                     breaks = breaks, 
                     key = TRUE, 
                     trace = "none", 
                     main = "Marker Correlation_DS_TabibSole", 
                     key.title = "Correlation",   
                     cexRow = 0.6, 
                     cexCol = 0.6,    
                     distfun = function(x) dist(x, method = "euclidean"),
                     hclust = function(x) hclust(x, method = "ward.D2"), 
                     symkey = FALSE, 
                     Rowv=dendrogram_row,
                     Colv=dendrogram_col,
                     density.info = c("none")) +
  custom_theme(base_size=20)

dev.off()

print(heatmap)


```
# Generating a correlation matrix of a panel of selected senescence genes - EP predicted cells
```{r}
# Subset the Seurat object based on scpred_prediction
filtered_seurat_EP <- subset(seurat_integrated_3D, subset = scpred_prediction == "EP3D")

# Extract gene counts for filtered cells
gene_counts <- filtered_seurat_EP@assays$RNA@data

# Define gene list 
gene_list <- c("CDKN1A", "CDKN2A", "IGFBP3", "SERPINE1", "IGF1", "CXCL12", "IL6", "IL8", "MMP3", "TP53", "IL1B")

# Check if gene names in gene_list are present in row names (genes)
missing_genes <- gene_list[!gene_list %in% rownames(gene_counts)]
if (length(missing_genes) > 0) {
  stop(paste("Genes missing in gene counts:", paste(missing_genes, collapse = ", ")))
}

# Filter gene counts based on the gene list
gene_counts <- gene_counts[gene_list, ]
gene_counts <- as.matrix(gene_counts)
# Transpose gene counts matrix
gene_counts_transposed <- t(gene_counts)
gene_counts_transposed <- as.matrix(gene_counts_transposed)

# Calculate correlation matrix
correlation_matrix <- cor(gene_counts_transposed)

# Print correlation matrix
print(correlation_matrix)

# Define breaks and color palette
breaks <- unique(c(seq(-1,-0.0001,length=50),seq(-0.00001,0.01,length=50), seq(0.011,1,length=50)))
my_palette <- colorRampPalette(c("blue","white", "white", "red"))(length(breaks)-1)

library("gplots")
png(filename = "../Output/TabibSole_Integrated/Markercorrelation_EP_tabibsole.png", width = 800, height = 600)  # Set appropriate width and height

heatmap <- heatmap.2(correlation_matrix, 
                     col = my_palette, 
                     breaks = breaks, 
                     key = TRUE, 
                     trace = "none", 
                     main = "Marker Correlation_EP_TabibSole", 
                     key.title = "Correlation",   
                     cexRow = 0.6, 
                     cexCol = 0.6,    
                     distfun = function(x) dist(x, method = "euclidean"),
                     hclust = function(x) hclust(x, method = "ward.D2"), 
                     symkey = FALSE, 
                     Rowv = dendrogram_row,
                     Colv = dendrogram_col,
                     dendrogram = "both",
                     density.info = c("none")) +
  custom_theme(base_size=20)

dev.off()

print(heatmap)
```
# Extract example correlation plots
## CDKN1A vs CDKN2A
```{r}
# Extract counts for CDKN1A and CDKN2A from Seurat object - DS only
CDKN1A_counts <- filtered_seurat@assays$RNA@data["CDKN1A", ]
CDKN2A_counts <- filtered_seurat@assays$RNA@data["CDKN2A", ]

# Combine counts into a data frame
counts_df <- data.frame(CDKN1A = CDKN1A_counts, CDKN2A = CDKN2A_counts)

# Print the first few rows of the data frame
print(head(counts_df))

# Create scatterplot
ggplot(counts_df, aes(x = CDKN1A, y = CDKN2A)) +
  geom_point() +
  labs(x = "CDKN1A counts", y = "CDKN2A counts", title = "Scatterplot of CDKN1A vs CDKN2A counts_EP")

# Compute the Pearson correlation coefficient
correlation <- cor(counts_df$CDKN1A, counts_df$CDKN2A)
print(correlation)

# Create scatterplot 
ggplot(counts_df, aes(x = CDKN1A, y = CDKN2A)) +
  geom_point() +
  labs(x = "CDKN1A counts", y = "CDKN2A counts", title = "Scatterplot of CDKN1A vs CDKN2A counts_DS") +
  annotate("text", x = min(counts_df$CDKN1A), y = max(counts_df$CDKN2A), 
           label = paste("Pearson Correlation:", round(correlation, 2)), 
           hjust = -1, vjust = 1, color = "black", size = 6) +
  custom_theme(base_size=20)

ggsave(
  "scatterplot_CDKN1ACDKN2A_DS.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/TabibSole_Integrated"
  )

## Run the same for EP
# Extract just EP
filtered_seurat_EP <- subset(seurat_integrated_3D, subset = scpred_prediction == "EP3D")

# Extract counts for CDKN1A and CDKN2A from Seurat object - DS only
CDKN1A_counts <- filtered_seurat_EP@assays$RNA@data["CDKN1A", ]
CDKN2A_counts <- filtered_seurat_EP@assays$RNA@data["CDKN2A", ]

# Combine counts into a data frame
counts_df <- data.frame(CDKN1A = CDKN1A_counts, CDKN2A = CDKN2A_counts)

# Print the first few rows of the data frame
print(head(counts_df))

# Create scatterplot
ggplot(counts_df, aes(x = CDKN1A, y = CDKN2A)) +
  geom_point() +
  labs(x = "CDKN1A counts", y = "CDKN2A counts", title = "Scatterplot of CDKN1A vs CDKN2A counts_EP")

# Compute the Pearson correlation coefficient
correlation <- cor(counts_df$CDKN1A, counts_df$CDKN2A)
print(correlation)

# Create scatterplot 
ggplot(counts_df, aes(x = CDKN1A, y = CDKN2A)) +
  geom_point() +
  labs(x = "CDKN1A counts", y = "CDKN2A counts", title = "Scatterplot of CDKN1A vs CDKN2A counts_EP") +
  annotate("text", x = min(counts_df$CDKN1A), y = max(counts_df$CDKN2A), 
           label = paste("Pearson Correlation:", round(correlation, 2)), 
           hjust = -1, vjust = 1, color = "black", size = 6) +
  custom_theme(base_size=20)

ggsave(
  "scatterplot_CDKN1ACDKN2A_EP.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/TabibSole_Integrated"
  )


```

##IL6 vs IL8
```{r}
# Extract counts for IL6 and IL8 from Seurat object - DS
IL6_counts <- filtered_seurat@assays$RNA@data["IL6", ]
IL8_counts <- filtered_seurat@assays$RNA@data["IL8", ]

# Combine counts into a data frame
counts_df <- data.frame(IL6 = IL6_counts, IL8 = IL8_counts)

# Print the first few rows of the data frame
print(head(counts_df))

# Create scatterplot
ggplot(counts_df, aes(x = IL6, y = IL8)) +
  geom_point() +
  labs(x = "IL6 counts", y = "IL8 counts", title = "Scatterplot of IL6 vs IL8 counts")

# Compute the Pearson correlation coefficient
correlation <- cor(counts_df$IL6, counts_df$IL8)
print(correlation)

# Create scatterplot 
ggplot(counts_df, aes(x = IL6, y = IL8)) +
  geom_point() +
  labs(x = "IL6 counts", y = "IL8 counts", title = "Scatterplot of IL6 vs IL8 counts_DS") +
  annotate("text", x = min(counts_df$IL6), y = max(counts_df$IL8), 
           label = paste("Pearson Correlation:", round(correlation, 2)), 
           hjust = -1, vjust = 1, color = "black", size = 6) +
  custom_theme(base_size=20)

ggsave(
  "scatterplot_IL6IL8_DS.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/TabibSole_Integrated"
  )

# Extract counts for IL6 and IL8 from Seurat object - EP
IL6_counts <- filtered_seurat_EP@assays$RNA@data["IL6", ]
IL8_counts <- filtered_seurat_EP@assays$RNA@data["IL8", ]

# Combine counts into a data frame
counts_df <- data.frame(IL6 = IL6_counts, IL8 = IL8_counts)

# Print the first few rows of the data frame
print(head(counts_df))

# Create scatterplot
ggplot(counts_df, aes(x = IL6, y = IL8)) +
  geom_point() +
  labs(x = "IL6 counts", y = "IL8 counts", title = "Scatterplot of IL6 vs IL8 counts")

# Compute the Pearson correlation coefficient
correlation <- cor(counts_df$IL6, counts_df$IL8)
print(correlation)

# Create scatterplot 
ggplot(counts_df, aes(x = IL6, y = IL8)) +
  geom_point() +
  labs(x = "IL6 counts", y = "IL8 counts", title = "Scatterplot of IL6 vs IL8 counts_EP") +
  annotate("text", x = min(counts_df$IL6), y = max(counts_df$IL8), 
           label = paste("Pearson Correlation:", round(correlation, 2)), 
           hjust = -1, vjust = 1, color = "black", size = 6) +
  custom_theme(base_size=20)

ggsave(
  "scatterplot_IL6IL8_EP.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/TabibSole_Integrated"
  )
```
##IGFBP3 vs TP53
```{r}
# Extract counts for IGFBP3 and TP53 from Seurat object - DS
IGFBP3_counts <- filtered_seurat@assays$RNA@data["IGFBP3", ]
TP53_counts <- filtered_seurat@assays$RNA@data["TP53", ]

# Combine counts into a data frame
counts_df <- data.frame(IGFBP3 = IGFBP3_counts, TP53 = TP53_counts)

# Print the first few rows of the data frame
print(head(counts_df))

# Create scatterplot
ggplot(counts_df, aes(x = IGFBP3, y = TP53)) +
  geom_point() +
  labs(x = "IGFBP3 counts", y = "TP53 counts", title = "Scatterplot of IGFBP3 vs TP53 counts_DS")

# Compute the Pearson correlation coefficient
correlation <- cor(counts_df$IGFBP3, counts_df$TP53)
print(correlation)

# Create scatterplot 
ggplot(counts_df, aes(x = IGFBP3, y = TP53)) +
  geom_point() +
  labs(x = "IGFBP3 counts", y = "TP53 counts", title = "Scatterplot of IGFBP3 vs TP53 counts_DS") +
  annotate("text", x = min(counts_df$IGFBP3), y = max(counts_df$TP53), 
           label = paste("Pearson Correlation:", round(correlation, 2)), 
           hjust = -1, vjust = 1, color = "black", size = 6) +
  custom_theme(base_size=20)

ggsave(
  "scatterplot_IGFBP3TP53_DS.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/TabibSole_Integrated"
  )

# Extract counts for IGFBP3 and TP53 from Seurat object - EP
IGFBP3_counts <- filtered_seurat_EP@assays$RNA@data["IGFBP3", ]
TP53_counts <- filtered_seurat_EP@assays$RNA@data["TP53", ]

# Combine counts into a data frame
counts_df <- data.frame(IGFBP3 = IGFBP3_counts, TP53 = TP53_counts)

# Print the first few rows of the data frame
print(head(counts_df))

# Create scatterplot
ggplot(counts_df, aes(x = IGFBP3, y = TP53)) +
  geom_point() +
  labs(x = "IGFBP3 counts", y = "TP53 counts", title = "Scatterplot of IGFBP3 vs TP53 counts_EP")

# Compute the Pearson correlation coefficient
correlation <- cor(counts_df$IGFBP3, counts_df$TP53)
print(correlation)

# Create scatterplot 
ggplot(counts_df, aes(x = IGFBP3, y = TP53)) +
  geom_point() +
  labs(x = "IGFBP3 counts", y = "TP53 counts", title = "Scatterplot of IGFBP3 vs TP53 counts_EP") +
  annotate("text", x = min(counts_df$IGFBP3), y = max(counts_df$TP53), 
           label = paste("Pearson Correlation:", round(correlation, 2)), 
           hjust = -1, vjust = 1, color = "black", size = 6) +
  custom_theme(base_size=20)

ggsave(
  "scatterplot_IGFBP3TP53_EP.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/TabibSole_Integrated"
  )
```

### Session information
```{r session_info_99}
sessionInfo()
```