---
title: "1_HDF_data"
author: "Bethany Hughes"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  rmarkdown::html_document:
    toc: TRUE
  
---

## Install/Load dependencies:
```{r load_packages}
library(dplyr)
library(Seurat)
library(patchwork)
library(readxl)
library(tidyseurat)
library(ggplot2)
library(tidyverse)
library(RCurl)
library(cowplot)
library(sctransform)
library(here)
library(stringr)
library(reshape2)
```

#Load input 10X data
```{r}
data_dir <- "../Input/filtered_feature_bc_matrix"

list.files(data_dir)
hdf.data <- Read10X(data.dir=data_dir, gene.column=2, cell.column=1, strip.suffix=TRUE)
```

## Inputting file which matches barcode identities with input sample identity (e.g. EP/DS/Mix) [All Data Transcript Counts_cell_annotations.csv](../Input/All Data Transcript Counts_cell_annotations.csv)
```{r}
cell_labels<-read.csv("../Input/All Data Transcript Counts Cleaned_suffixes.csv",header=FALSE)
cell_labels<-as.data.frame(t(cell_labels))
cell_labels<-cell_labels[!duplicated(cell_labels[,c('V2')]),]
rownames(cell_labels)<-cell_labels$V2
colnames(cell_labels)<-c("Sample_Source","Barcode")
```

## Merging sample identity data with barcodes
```{r}
rawUMI_cols<-as.data.frame(colnames(hdf.data))
colnames(rawUMI_cols)[1]<-"Barcode"
cells_with_metadata<-merge(rawUMI_cols,cell_labels,by="Barcode")
```

## Creating Seurat object
```{r}
hdfseurat <- CreateSeuratObject(counts = hdf.data, project = "MachineLearningSenescence", min.cells=3, min.features=200)
hdfseurat

#Adding percentage of mitochondrial genes
hdfseurat[["percent.mt"]] <- PercentageFeatureSet(hdfseurat, pattern = "^MT-")
metadata<-hdfseurat@meta.data

#Adding a barcode column
metadata$Barcode<-rownames(metadata)
```

## Matching sample type with barcode in the Seurat metadata
```{r}
DS2D_Names <- metadata %>% 
  filter(((str_detect(Barcode, "1")))) %>%
  select(Barcode) %>%
  mutate(Sample = "DS2D") 
DS3D_Names <- metadata %>% 
  filter(((str_detect(Barcode, "2")))) %>%
  select(Barcode) %>%
  mutate(Sample = "DS3D") 
EP2D_Names <- metadata %>% 
  filter(((str_detect(Barcode, "3")))) %>%
  select(Barcode) %>%
  mutate(Sample = "EP2D")
EP3D_Names <- metadata %>% 
  filter(((str_detect(Barcode, "4")))) %>%
  select(Barcode) %>%
  mutate(Sample = "EP3D")
Mix2D_Names <- metadata %>% 
  filter(((str_detect(Barcode, "5")))) %>%
  select(Barcode) %>%
  mutate(Sample = "Mix2D")
Mix3D_Names <- metadata %>% 
  filter(((str_detect(Barcode, "6")))) %>%
  select(Barcode) %>%
  mutate(Sample = "Mix3D")


x<-rbind(DS2D_Names,DS3D_Names,EP2D_Names,EP3D_Names,Mix2D_Names,Mix3D_Names)

metadata_merge<-merge(metadata,x,by="Barcode")
rownames(metadata_merge)<-metadata_merge$Barcode

#Recreate Seurat object with new metadata, including genes which are expressed in at least 3 cells, and cells with more than 200 genes sequenced
hdfseurat <- CreateSeuratObject(counts = hdf.data, project = "MachineLearningSenescence", min.cells=3, min.features=200, meta.data = metadata_merge)   

```


## Subsetting mix cells - to be removed once original files are fixed
```{r}
#Subsetting Mix cells - this will be excluded when I sort initial file
hdfseurat <- subset(hdfseurat, Sample == "DS2D" | Sample == "EP2D" | Sample == "DS3D" | Sample == "EP3D")
```

### Session information
```{r session_info_99}
sessionInfo()
```

This document was processed on: `r Sys.Date()`.