---
title: "20_lungs"
author: "BH"
date: "2024-02-23"
output:
 pdf_document:
    toc: yes
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load_packages}
library("dplyr")
library("Seurat")
library("patchwork")
library("readxl")
library("tidyseurat")
library("ggplot2")
library("tidyverse")
library("RCurl")
library("cowplot")
library("sctransform")
library("scPred")
library("Seurat")
library("magrittr")
library("caret")
library("dplyr")
library("janitor")
library("factoextra")
#library("EnsDb.Hsapiens.v86")



options(future.globals.maxSize = 12000 * 1024^2)

```
# To test SenPred on fibroblasts from another tissue, the lung, I extracted data from the human cell atlas using the following paper:
Sikkema, L., Ramírez-Suástegui, C., Strobl, D.C., Gillett, T.E., Zappia, L., Madissoon, E., Markov, N.S., Zaragosi, L.E., Ji, Y., Ansari, M. and Arguel, M.J., 2023. An integrated cell atlas of the lung in health and disease. Nature Medicine, 29(6), pp.1563-1577.

Data was downloaded from cellxgene at the following link:
https://cellxgene.cziscience.com/collections/6f6d381a-7701-4781-935c-db10d30de293

```{r}
custom_theme <- readRDS(file = "custom_theme.rds")
hdf_mda2D <- readRDS(file = "hdf_mda2D.rds")
hdf_mda3D <- readRDS(file="hdf_mda3D.rds")

lung <- readRDS("../Inputfiles/lungatlas.rds")

unique_studies <- unique(lung@meta.data$dataset)
unique_terms <- unique(lung@meta.data$scanvi_label)
unique_terms

#First subsetted for cells that were pre-defined as from 'fibroblast lineage'
lung_fibroblast <- subset(lung, subset = scanvi_label == 'Fibroblast lineage')

```

```{r}
unique_terms2 <- unique (lung_fibroblast@meta.data$ann_finest_level)
unique_terms2 <- as.data.frame(unique_terms2)

#Then, subset for actual fibroblasts
lung_fibroblast2 <- subset(lung_fibroblast, subset = ann_finest_level %in% c("Adventitial fibroblasts", "Peribronchial fibroblasts", "Alveolar fibroblasts", "Myofibroblasts", "Subpleural fibroblasts"))

#Subset to only use data from healthy lung condition donors
lung_fibroblast3 <- subset(lung_fibroblast2, subset = lung_condition == "Healthy")

ages <- unique(lung_fibroblast3@meta.data$age_or_mean_of_age_range)
ages <- as.data.frame(ages)

#Remove any donors where the age was not reported
lung_fibroblast4 <- subset(lung_fibroblast3, subset = age_or_mean_of_age_range != "NaN")

ages2 <- unique(lung_fibroblast4@meta.data$age_or_mean_of_age_range)
ages2 <- as.data.frame(ages2)
 
#Only select donors with more than 100 cells
# Create a vector of ages to filter
selected_ages <- c("25", "29", "30", "32.5", "34", "35", "42.5", "49",
                   "50", "52", "52.5", "55", "57", "57.5", "59", "62",
                   "64", "67.5")

# Filter the Seurat object based on the selected ages
lung_fibroblast4 <- subset(lung_fibroblast4, 
                                    subset = age_or_mean_of_age_range %in% selected_ages)
```

## Converting Ensembl ID to gene_symbol
```{r}
#install.packages("EnsDb.Hsapiens.v86")
library(EnsDb.Hsapiens.v86)

Gene_ID <- lung_fibroblast4@assays[["RNA"]]@counts@Dimnames[[1]]
Gene_ID <- as.data.frame(Gene_ID)

gene_name <- mapIds (EnsDb.Hsapiens.v86,
                     keys = Gene_ID$Gene_ID,
                     keytype = 'GENEID',
                     column = 'SYMBOL')

gene_name <- as.data.frame(gene_name)

# Function to add '-' to gene names starting with 'MT'
add_hyphen <- function(x) {
  # Replace 'MT' with 'MT-' for gene names starting with 'MT'
  x <- gsub("^MT", "MT-", x)
  return(x)
}

# Apply the function to the 'gene_name' column
gene_name$gene_name <- add_hyphen(gene_name$gene_name)


gene_name$gene_ids <- Gene_ID [["Gene_ID"]]

# Combine gene_name with Gene_ID
combined_data <- merge(Gene_ID, gene_name, by.x = "Gene_ID", by.y = "gene_ids", all.x = TRUE)

library(dplyr)

# Assuming combined_data is your DataFrame
duplicate_rows <- combined_data %>%
  dplyr::filter(duplicated(gene_name) | duplicated(gene_name, fromLast = TRUE))


# Assuming combined_data is your DataFrame
combined_data <- combined_data %>%
  mutate(gene_name = ifelse(duplicated(gene_name) | duplicated(gene_name, fromLast = TRUE),
                            paste0(gene_name, '-1'),
                            gene_name))

MT_genes <- combined_data %>%
  dplyr::filter(startsWith(gene_name, "MT-"))



# Assuming gene_name is your DataFrame containing gene symbols
genes_with_double_hyphen <- gene_name %>%
  dplyr::filter(grepl("--", gene_name))

# Assuming combined_data is your DataFrame containing the gene_ids column
gene_ids <- genes_with_double_hyphen$gene_ids
gene_ids <- as.data.frame(gene_ids)
gene_ids <- gene_ids %>%
  mutate(gene_ids = str_replace(gene_ids, "--", "-"))
gene_ids_mito <- gene_ids

# Save the gene_ids column as a CSV file
write.csv(gene_ids, file = "../Output/LungAtlas/mito_gene_ids.csv", row.names = FALSE)

```


```{r}

pattern <- gene_ids_mito$gene_ids

# Calculate the percentage of cells expressing the features in gene_ids_mito
lung_fibroblast4[["percent.mt"]] <- PercentageFeatureSet(lung_fibroblast4, pattern = pattern)



#Adding no. of genes per UMI.
lung_fibroblast4$log10GenesPerUMI <- log10(lung_fibroblast4$nFeature_RNA) / log10(lung_fibroblast4$nCount_RNA)

#Visualise the number of cells per age_or_mean_of_age_range
CellNumbers <- lung_fibroblast4@meta.data %>% 
  	           ggplot(aes(x=age_or_mean_of_age_range, fill=age_or_mean_of_age_range)) + 
  	           geom_bar() +
  	           theme_classic() +
  	           theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	           theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	           ggtitle("NCells")

CellNumbers

#Save plot
ggsave(
  "NumbersPerage_or_mean_of_age_range.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/LungAtlas"
  )

```
## Visualising QC metrics - nFeature_RNA, nCount_RNA, percent.mt
```{r}
# Create plot
violin <- VlnPlot(lung_fibroblast4, features =c("nFeature_RNA", "nCount_RNA", "percent.mt"))
violin

#Save plot
ggsave(
  "HDF_QC_violin.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/LungAtlas"
  )
```

## Visualising the number of UMIs per cell (nCount_RNA)
```{r}
#Number of UMIs (transcripts) per cell
UMIspercell <- lung_fibroblast4@meta.data %>% 
  	           ggplot(aes(color=age_or_mean_of_age_range, x=nCount_RNA, fill= age_or_mean_of_age_range)) + 
  	           geom_density(alpha = 0.2) + 
  	           scale_x_log10() + 
  	           theme_classic() +
  	           ylab("Cell density") +
  	           geom_vline(xintercept = 1000) 

UMIspercell

#Save plot
ggsave(
  "UMI-per-cell.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/LungAtlas"
  )

```
## Visualising number of genes per cell (nFeature_RNA)
```{r}
# Visualize the distribution of genes detected per cell via histogram
hist_feature<- lung_fibroblast4@meta.data %>% 
  	           ggplot(aes(color=age_or_mean_of_age_range, x=nFeature_RNA, fill= age_or_mean_of_age_range)) + 
  	           geom_density(alpha = 0.2) + 
  	           theme_classic() +
  	           scale_x_log10() + 
  	           geom_vline(xintercept = c(400)) 

hist_feature

#Save plot
ggsave(
  "hist_featureRNA.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/LungAtlas"
  )

# Visualize the distribution of genes detected per cell via boxplot
boxplot_feature <- lung_fibroblast4@meta.data %>% 
    	             ggplot(aes(x=age_or_mean_of_age_range, y=log10(nFeature_RNA), fill=age_or_mean_of_age_range)) + 
  	               geom_boxplot() + 
  	               theme_classic() +
  	               theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	               theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	               ggtitle("NCells vs NGenes")

boxplot_feature

#Save plot
ggsave(
  "boxplot_featureRNA.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/LungAtlas"
  )
    
```

## Combining number of genes, UMIs and mitochondrial percentage, to show thresholding boundaries
```{r}
thresholding <- lung_fibroblast4@meta.data %>% 
  	            ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) + 
  	            geom_point() + 
	              scale_colour_gradient(low = "gray90", high = "black") +
  	            stat_smooth(method=lm) +
  	            scale_x_log10() + 
  	            scale_y_log10() + 
  	            theme_classic() +
  	            geom_vline(xintercept = 1000) +
  	            geom_hline(yintercept = 400) +
  	            facet_wrap(~age_or_mean_of_age_range) 

thresholding

ggsave(
  "thresholding_feature_count_mito.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/LungAtlas"
  )
```
## Mitochondrial gene expression per cell
```{r}
# Visualize the distribution of mitochondrial gene expression detected per cell
mitochondrial_genes <- lung_fibroblast4@meta.data %>% 
                    	 ggplot(aes(color=age_or_mean_of_age_range, x=percent.mt, fill=age_or_mean_of_age_range)) + 
  	                   geom_density(alpha = 0.2) + 
  	                   scale_x_log10() + 
  	                   theme_classic() +
  	                   geom_vline(xintercept = 12) 

mitochondrial_genes

ggsave(
  "mitochondrial_hist.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/LungAtlas"
  )
```
## Visualising the proportion of genes per UMI
```{r}
genesperUMI <- lung_fibroblast4@meta.data %>%
  	           ggplot(aes(x=log10GenesPerUMI, color = age_or_mean_of_age_range, fill=age_or_mean_of_age_range)) +
  	           geom_density(alpha = 0.2) +
  	           theme_classic() +
  	           geom_vline(xintercept = 0.8) 

genesperUMI

ggsave(
  "genesperUMI.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/LungAtlas"
  )
```

#Subsetting seurat object
```{r}
lung_fibroblast4_filt <- subset(x=lung_fibroblast4, 
                         subset= (percent.mt <=12) &
                            (nFeature_RNA >=400) &
                            (nCount_RNA >= 1000))

#Visualising number of cells before and after subsetting
df.list <- list(Original=lung_fibroblast4@meta.data, Filtered=lung_fibroblast4_filt@meta.data)
dat <- stack(lapply(df.list, `[[`, "age_or_mean_of_age_range"))
numbers_after_filt <- ggplot(dat, aes(x=values, fill=ind)) + geom_bar(position="dodge") 

numbers_after_filt

ggsave(
  "barplot_beforeandafterfilt.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/LungAtlas"
  )

```

## Normalisation - Normalise gene expression per cell by the total gene expression within that cell. This is then multiplied by a default scale factor (10,000), and log-transformed.
```{r}

lung_fibroblast4_filt <- NormalizeData(lung_fibroblast4_filt)
ages <- unique(lung_fibroblast4_filt@meta.data$age_or_mean_of_age_range)
ages <- as.data.frame(ages)

```


```{r}
library(EnsDb.Hsapiens.v86)

Gene_ID <- lung_fibroblast4_filt@assays[["RNA"]]@counts@Dimnames[[1]]
Gene_ID <- as.data.frame(Gene_ID)

gene_name <- mapIds (EnsDb.Hsapiens.v86,
                     keys = Gene_ID$Gene_ID,
                     keytype = 'GENEID',
                     column = 'SYMBOL')

gene_name <- as.data.frame(gene_name)

# Assuming gene_name is your data frame
# Assuming gene_name is your data frame
gene_name <- gene_name %>%
  mutate(gene_name = ifelse(duplicated(gene_name) | duplicated(gene_name, fromLast = TRUE), 
                            ifelse(duplicated(gene_name) & !duplicated(gene_name, fromLast = TRUE), 
                                   paste0(gene_name, "-1"), 
                                   gene_name), 
                            gene_name))
gene_name$gene_ids <- Gene_ID[["Gene_ID"]]
lung_fibroblast4_filt@assays[["RNA"]]@counts@Dimnames[[1]] <- gene_name[["gene_name"]]

lung_fibroblast4_filt@assays[["RNA"]]@data@Dimnames[[1]] <- gene_name[["gene_name"]]

rownames(lung_fibroblast4_filt@assays$RNA@meta.features) <- gene_name[["gene_name"]]

saveRDS(lung_fibroblast4_filt, file = "../Output/lung_fibroblast4_filt.rds")
```




### Session information
```{r session_info_99}
sessionInfo()
```

This document was processed on: `r Sys.Date()`.
