---
title: "3_HDF_2DCells"
author: "Beth"
date: "24/09/2021"
output:
 pdf_document:
    toc: yes
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```
## Install/Load dependencies:
```{r load_packages, results = FALSE}
library(dplyr)
library(Seurat)
library(patchwork)
library(readxl)
library(tidyseurat)
library(ggplot2)
library(tidyverse)
library(RCurl)
library(cowplot)
library(sctransform)
library(here)
library(stringr)
library(reshape2)
library(ggplot2)
```

# Filtering Seurat to select just 2D cells for clustering 
```{r}
hdfseurat_filt <- readRDS(file = "hdfseurat_filt.rds")
custom_theme <- readRDS(file = "custom_theme.rds")


hdfseurat_filt_2D <- subset(hdfseurat_filt, Sample == "DS2D" | Sample == "EP2D")
```

# Plotting top 10 variable features in the 2D cells
```{r}
#Selecting highly variable features - high cell to cell variation
hdfseurat_filt_2D <- FindVariableFeatures(hdfseurat_filt_2D, selection.method ="vst", nfeatures=2000)
#Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(hdfseurat_filt_2D), 10)
top10
plot3 <- VariableFeaturePlot(hdfseurat_filt_2D)
#Plot variable features with labels
plot4 <- LabelPoints(plot=plot3, points = top10, repel=TRUE)
plot4

ggsave(
  "HDF_2D_variablefeatures.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/HDF_Data_Analysis"
  )
```
# Scaling the data before dimensionality reduction. Each gene is centred to a mean of 0, and scaled by the standard deviation of each gene.
```{r}
all.genes <- rownames(hdfseurat_filt_2D)
hdfseurat_filt_2D <- ScaleData(hdfseurat_filt_2D, features = all.genes)

#Performing a PCA
hdfseurat_filt_2D <- RunPCA(hdfseurat_filt_2D, features=VariableFeatures(object = hdfseurat_filt_2D))

#Plotting the PCA
PCA_no_label <- DimPlot(hdfseurat_filt_2D, reduction = "pca", seed = 123, combine=TRUE) + NoLegend()
PCA_Sample_label <- DimPlot(hdfseurat_filt_2D, reduction = "pca", seed = 123, group.by = "Sample")

PCA_no_label
PCA_Sample_label
```


## Heatmaps to investigate the genes influencing the top PCs
```{r}
DimHeatmap(hdfseurat_filt_2D, dims=1, cells=500, balanced=TRUE)
DimHeatmap(hdfseurat_filt_2D, dims=1:15, cells=500, balanced=TRUE)
ElbowPlot(hdfseurat_filt_2D) 

#Suggests top 10 PCs explain most of the variance. Test this with jackstraw plot
#hdfseurat_jackstraw <- JackStraw(hdfseurat_filt_2D, num.replicate = 100)
#hdfseurat_jackstraw <- ScoreJackStraw(hdfseurat_jackstraw, dims = 1:20)
#JackStrawPlot(hdfseurat_jackstraw, dims = 1:15)
```

## Clustering the cells using KNN
```{r}
#Use K-nearest neighbours based on PCA distance in FindNeighbours() function based on previously definited dimensions (10) from elbow plot
hdfseurat_filt_2D <- FindNeighbors(hdfseurat_filt_2D, dims =1:10)

#group cells using findclusters function
hdfseurat_filt_2D <- FindClusters(hdfseurat_filt_2D, resolution = 0.4)

#Look at cluster ID's of first 5 cells
head(Idents(hdfseurat_filt_2D), 5)
```

## Plot UMAP
```{r}
#Running a UMAP
hdfseurat_filt_2D <- RunUMAP(hdfseurat_filt_2D, dims=1:10)


#Plotting UMAP with cluster IDs
DimPlot(hdfseurat_filt_2D, reduction = "umap", seed = 123, cols = c("#9b5fe0", "#16a5d8", "#60dbe8", "#8bd346", "#efdf48", "#f9a52c", "#d64e12" )) + 
  custom_theme(base_size=20)
ggsave(
  "UMAP_HDF_2D.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/HDF_Data_Analysis"
  )

#Colouring above UMAP plot with Sample type
DimPlot(hdfseurat_filt_2D, reduction = "umap",group.by = "Sample", seed=123, cols = c( "#24B3A8", "#FF92A5")) +
  custom_theme(base_size=20)
ggsave(
  "UMAP_HDF_2D_samplecolour.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/HDF_Data_Analysis"
  )
```
#Finding DEGs/cluster biomarkers
```{r}
#Find markers for every cluster compared to all remaining cells, report only the positive ones
hdfseurat.markers <- FindAllMarkers(hdfseurat_filt_2D, only.pos=FALSE, min.pct=0.25, logfc.threshold=0.25)

write.csv(hdfseurat.markers,"../Output/HDF_Data_Analysis/HDF2D_clustermarkers_posandneg.csv")

hdfseurat.markers %>%
  group_by(cluster)%>%
  top_n(n=2, wt = avg_log2FC)
```
## Saving Seurat Object with cluster identities
```{r}
saveRDS(hdfseurat_filt_2D, file = "../Output/Seurat_Objects/hdfseuratfiltered_2Dcells.rds")
```

### Session information
```{r, echo=FALSE}
sessionInfo()
```

This document was processed on: `r Sys.Date()`.