---
title: "4_ScPred_2D"
author: "BH"
date: '2022-03-23'
output: html_document
---
## Building machine learning models of proliferative and senescent HDFs in 2D, using ScPred package (see https://github.com/powellgenomicslab/scPred)

## Loading libraries
```{r}
library("scPred")
library("Seurat")
library("magrittr")
library("caret")
library("dplyr")
library("janitor")
library("factoextra")
library("MLmetrics")
```

## Seperating dataset into training and testing. 80% training, and 20% testing
```{r}
new_col <- 1:2193

hdfseurat_filt_2D@meta.data <- cbind(hdfseurat_filt_2D@meta.data, new_col)

hdfseurat_filt_2D@meta.data$new_col <- 1:nrow(hdfseurat_filt_2D@meta.data)
train <- hdfseurat_filt_2D@meta.data %>% dplyr::sample_frac(.8)
test  <- dplyr::anti_join(hdfseurat_filt_2D@meta.data, train, by = 'new_col')

test <- test$new_col
train <- train$new_col

hdfseurat_filt_2D@meta.data$new_col[hdfseurat_filt_2D@meta.data$new_col %in% test] <- "Test"
hdfseurat_filt_2D@meta.data$new_col[hdfseurat_filt_2D@meta.data$new_col %in% train] <- "Train"

obj_list <- SplitObject(hdfseurat_filt_2D, split.by = "new_col")
```

## Normalising, scaling and running PCA on training and testing data
```{r}
training <- obj_list$Train %>% 
  NormalizeData() %>% 
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA() %>% 
  RunUMAP(dims = 1:10)

DimPlot(training, group.by = "Sample", label = FALSE, repel = TRUE, seed =123)

ggsave(
  "UMAP_2D_training.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/HDF_ScPred"
  )

testing <- obj_list$Test%>%
  NormalizeData() %>%
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA() %>% 
  RunUMAP(dims = 1:10)

DimPlot(testing, group.by = "Sample", label = FALSE, repel=TRUE, seed = 123)

ggsave(
  "UMAP_2D_testing.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/HDF_ScPred"
  )

##Save training and testing datasets
saveRDS(training, file = "../Output/Seurat_Objects/hdfseuratfiltered_2Dtraining.rds")
saveRDS(testing, file =  "../Output/Seurat_Objects/hdfseuratfiltered_2Dtesting.rds" )

## Note - every time you run the script you will get slightly different results. This is because the training and testing splits are random. If you want to exactly replicate your results, save the training and testing files and load them in each time.

#training <- readRDS("../Output/Seurat_Objects/hdfseuratfiltered_2Dtraining.rds")
#testing <- readRDS("../Output/Seurat_Objects/hdfseuratfiltered_2Dtesting.rds")

```

```{r}
#Using the training dataset to create and evaluate different models - no PC filters

hdfseurat_scPred <- getFeatureSpace(training, "Sample")


#This model is the svm Radial model
hdfseurat_svmradial2D <- trainModel(hdfseurat_scPred)
get_probabilities(hdfseurat_svmradial2D) %>% head()
get_scpred(hdfseurat_svmradial2D)
hdfseurat_svmradial2D
get_classifiers(hdfseurat_svmradial2D)


#This model is the mda model
hdf_mda2D <- trainModel(hdfseurat_scPred, model = "mda")
get_probabilities(hdf_mda2D) %>% head()
get_scpred(hdf_mda2D)
hdf_mda2D

#This model is logistic regression
hdf_glm2D <- trainModel(hdfseurat_scPred, model = "glm")
get_probabilities(hdf_glm2D)%>% head()
get_scpred(hdf_glm2D)
hdf_glm2D

#This model is knn
hdf_knn2D <- trainModel(hdfseurat_scPred, model = "knn")
get_probabilities(hdf_knn2D)%>% head()
get_scpred(hdf_knn2D)
hdf_knn2D
```

## Decided to take SVM and MDA forward for further analysis due to increased sens and spec. 
Testing on test dataset

```{r}

#SVM testing 
query <- NormalizeData(testing)
query <- scPredict(query, hdfseurat_svmradial2D, recompute_alignment = FALSE)

#Plotting UMAP of predicted senescent cells
query <- RunUMAP(query, reduction = "scpred", dims = 1:10)
DimPlot(query, group.by = "scpred_prediction", label = FALSE, repel = TRUE, seed = 123, cols = c("#1e2f97", "#AA336A", "antiquewhite4"))
ggsave(
  "UMAP_scpred_testing2D_SVM.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/HDF_Data_ScPred/2D_SVM"
  )
DimPlot(query, group.by = "Sample", label = FALSE, repel = TRUE, seed = 123, cols = c( "#24B3A8", "#FF92A5"))
ggsave(
  "UMAP_scpred_testing2D_SAMPLE.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/HDF_Data_ScPred/2D_SVM"
  )
FeaturePlot(query, c("scpred_DS2D", "scpred_EP2D"))
```
#Confusion matrix plot
```{r}
query_df <- query@meta.data
query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="DS2D" & query_df$scpred_prediction=="DS2D" ~ "DS2DT"))

ds2dt <- nrow(query_df[!is.na(query_df$Status),])

query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="DS2D" & query_df$scpred_prediction=="EP2D" ~ "EP2DF"))

ep2df <- nrow(query_df[!is.na(query_df$Status),])

query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="EP2D" & query_df$scpred_prediction=="EP2D" ~ "EP2DT"))

ep2dt <- nrow(query_df[!is.na(query_df$Status),])


query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="EP2D" & query_df$scpred_prediction=="DS2D" ~ "DS2DF"))

ds2df <- nrow(query_df[!is.na(query_df$Status),])

unassigned <- length(which(query_df$scpred_prediction == "unassigned"))

#Plotting into confusion matrix

table <- matrix(c(ds2dt, ep2df, ds2df, ep2dt),ncol=2,byrow=TRUE)
colnames(table) <- c("DS", "EP")
rownames(table) <- c("DS", "EP")
table <- as.table(table)
table

perc <- prop.table(table, 1)*100
perc

perc <- round(perc, digits=2)

cm <- confusionMatrix(perc)


# Required library for saving plots
library(png)

# Drawing confusion matrix into visual table and save as a PNG file
draw_confusion_matrix <- function(cm, output_file) {
  # Create a function to draw the matrix plot
  draw_matrix_plot <- function() {
    layout(matrix(c(1, 1, 2)))
    par(mar = c(2, 2, 2, 2))
    plot(c(100, 345), c(300, 450), type = "n", xlab = "", ylab = "", xaxt = 'n', yaxt = 'n')
    title('SVM Confusion Matrix (%)', cex.main = 2)

    # create the matrix
    rect(150, 430, 240, 370, col = '#32CD32')
    text(195, 440, 'DS', cex = 1.2)
    rect(250, 430, 340, 370, col = '#FF4500')
    text(295, 440, 'EP', cex = 1.2)
    text(125, 370, 'Predicted', cex = 1.3, srt = 90, font = 2)
    text(245, 450, 'Actual', cex = 1.3, font = 2)
    rect(150, 305, 240, 365, col = '#FF4500')
    rect(250, 305, 340, 365, col = '#32CD32')
    text(140, 400, 'DS', cex = 1.2, srt = 90)
    text(140, 335, 'EP', cex = 1.2, srt = 90)

    # add in the cm results
    res <- as.numeric(cm$table)
    text(195, 400, res[1], cex = 1.6, font = 2, col = 'white')
    text(195, 335, res[2], cex = 1.6, font = 2, col = 'white')
    text(295, 400, res[3], cex = 1.6, font = 2, col = 'white')
    text(295, 335, res[4], cex = 1.6, font = 2, col = 'white')
    text(110, 300, 'Unassigned =', cex = 0.9, font = 1, col = 'black')
    text(130, 300, unassigned, cex = 0.9, font = 1, col = 'black')
  }

  # Call the function to draw the matrix plot
  draw_matrix_plot()

  # Save the plot as a PNG file
  png(filename = output_file, width = 800, height = 600)  # Set appropriate width and height
  draw_matrix_plot()
  dev.off()  # Turn off the PNG device
}


your_confusion_matrix <- cm
output_file_path <- "../Output/HDF_ScPred/CM2D_SVM_testing.png"
draw_confusion_matrix(your_confusion_matrix, output_file_path)


```
#Testing MDA model
```{r}
#SVM testing 
query <- NormalizeData(testing)
query <- scPredict(query, hdf_mda2D, recompute_alignment = FALSE)

#Plotting UMAP of predicted senescent cells
query <- RunUMAP(query, reduction = "scpred", dims = 1:10)
DimPlot(query, group.by = "scpred_prediction", label = FALSE, repel = TRUE, seed = 123, cols = c("darkblue", "deeppink3"))
ggsave(
  "UMAP_scpred_testing2D.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/HDF_Data_ScPred/2D_MDA"
  )
DimPlot(query, group.by = "Sample", label = FALSE, repel = TRUE, seed = 123, cols = c( "#24B3A8", "#FF92A5"))
ggsave(
  "UMAP_scpred_testing2D_SAMPLE.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/HDF_Data_ScPred/2D_MDA"
  )
FeaturePlot(query, c("scpred_DS2D", "scpred_EP2D"))
```
#Confusion matrix plot
```{r}
query_df <- query@meta.data
query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="DS2D" & query_df$scpred_prediction=="DS2D" ~ "DS2DT"))

ds2dt <- nrow(query_df[!is.na(query_df$Status),])

query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="DS2D" & query_df$scpred_prediction=="EP2D" ~ "EP2DF"))

ep2df <- nrow(query_df[!is.na(query_df$Status),])

query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="EP2D" & query_df$scpred_prediction=="EP2D" ~ "EP2DT"))

ep2dt <- nrow(query_df[!is.na(query_df$Status),])


query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="EP2D" & query_df$scpred_prediction=="DS2D" ~ "DS2DF"))

ds2df <- nrow(query_df[!is.na(query_df$Status),])

unassigned <- length(which(query_df$scpred_prediction == "unassigned"))

#Plotting into confusion matrix

table <- matrix(c(ds2dt, ep2df, ds2df, ep2dt),ncol=2,byrow=TRUE)
colnames(table) <- c("DS", "EP")
rownames(table) <- c("DS", "EP")
table <- as.table(table)
table

perc <- prop.table(table, 1)*100
perc

perc <- round(perc, digits=2)

cm <- confusionMatrix(perc)

# Required library for saving plots
library(png)

# Drawing confusion matrix into visual table and save as a PNG file
draw_confusion_matrix <- function(cm, output_file) {
  # Create a function to draw the matrix plot
  draw_matrix_plot <- function() {
    layout(matrix(c(1, 1, 2)))
    par(mar = c(2, 2, 2, 2))
    plot(c(100, 345), c(300, 450), type = "n", xlab = "", ylab = "", xaxt = 'n', yaxt = 'n')
    title('MDA Confusion Matrix (%)', cex.main = 2)

    # create the matrix
    rect(150, 430, 240, 370, col = '#32CD32')
    text(195, 440, 'DS', cex = 1.2)
    rect(250, 430, 340, 370, col = '#FF4500')
    text(295, 440, 'EP', cex = 1.2)
    text(125, 370, 'Predicted', cex = 1.3, srt = 90, font = 2)
    text(245, 450, 'Actual', cex = 1.3, font = 2)
    rect(150, 305, 240, 365, col = '#FF4500')
    rect(250, 305, 340, 365, col = '#32CD32')
    text(140, 400, 'DS', cex = 1.2, srt = 90)
    text(140, 335, 'EP', cex = 1.2, srt = 90)

    # add in the cm results
    res <- as.numeric(cm$table)
    text(195, 400, res[1], cex = 1.6, font = 2, col = 'white')
    text(195, 335, res[2], cex = 1.6, font = 2, col = 'white')
    text(295, 400, res[3], cex = 1.6, font = 2, col = 'white')
    text(295, 335, res[4], cex = 1.6, font = 2, col = 'white')
    text(110, 300, 'Unassigned =', cex = 0.9, font = 1, col = 'black')
    text(130, 300, unassigned, cex = 0.9, font = 1, col = 'black')
  }

  # Call the function to draw the matrix plot
  draw_matrix_plot()

  # Save the plot as a PNG file
  png(filename = output_file, width = 800, height = 600)  # Set appropriate width and height
  draw_matrix_plot()
  dev.off()  # Turn off the PNG device
}


your_confusion_matrix <- cm
output_file_path <- "../Output/HDF_ScPred/CM2D_MDA_testing.png"
draw_confusion_matrix(your_confusion_matrix, output_file_path)

```

### Session information
```{r session_info_99}
sessionInfo()
```

This document was processed on: `r Sys.Date()`.