---
title: "6_ScPred2D_Chan_EPESDS"
author: "BH"
date: '2022-11-15'
output:
  prettydoc::html_pretty:
    theme: cayman
    toc: yes
---
```{r load_packages}
library("dplyr")
library("Seurat")
library("patchwork")
library("readxl")
library("tidyseurat")
library("ggplot2")
library("tidyverse")
library("RCurl")
library("cowplot")
library("sctransform")
library("scPred")
library("Seurat")
library("magrittr")
library("caret")
library("dplyr")
library("janitor")
library("factoextra")

options(future.globals.maxSize = 12000 * 1024^2)

```

# Building 'Early Senescent' PDL50 cells into the model as an additional classifier

```{r}
#Subsetting PDL50 'early senescent' cells
ES <- subset(Chan_seurat2, PDL == "PDL_50")

#Adding PDL to a 'Sample' column to help with merge 
ESMeta = ES@meta.data
ESMeta["Sample"] <- "PDL50"
head(ESMeta)
ESMetaTrim <- subset(ESMeta, select = c("Sample"))
ES <- AddMetaData(ES, ESMetaTrim)

#Renaming barcode to 'Barcode' to match
ES$Barcode <- ES$barcode
ES$barcode <- NULL

#Normalising ES data
ES <- NormalizeData(ES, verbose = FALSE)

#Finding variable features in ES data
ES <- FindVariableFeatures(ES, selection.method = "vst", nfeatures = 2000)


hdfseurat_filt_2D <- NormalizeData(hdfseurat_filt_2D, verbose = FALSE)
hdfseurat_filt_2D <- FindVariableFeatures(hdfseurat_filt_2D, selection.method = "vst", nfeatures = 2000)

#Integrating two datasets by finding anchors
Anchors <- FindIntegrationAnchors(object.list = list(ES, hdfseurat_filt_2D), dims = 1:20)
Combined <- IntegrateData(anchorset = Anchors, dims = 1:20)
```

## Plotting top 10 variable features in the cells
```{r}
#Selecting highly variable features - high cell to cell variation
Combined<- FindVariableFeatures(Combined, selection.method ="vst", nfeatures=2000)
#Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(Combined), 10)
top10
plot3 <- VariableFeaturePlot(Combined)
#Plot variable features with labels
plot4 <- LabelPoints(plot=plot3, points = top10, repel=TRUE)
plot3
plot4

ggsave(
  "HDF_Chan_variablefeatures.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Early_senescence"
  )

```

## PCA 
```{r}
all.genes <- rownames(Combined)
HDF_Chan_combined <- ScaleData(Combined, features = all.genes)

#Performing a PCA
HDF_Chan_combined <- RunPCA(HDF_Chan_combined, features=VariableFeatures(object = HDF_Chan_combined))
DimPlot(HDF_Chan_combined, reduction = "pca")
ggsave(
  "HDF_Chan_PCA.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Early_senescence"
  )
```

## PCA coloured by sample type
```{r}

DimPlot(HDF_Chan_combined, reduction = "pca", group.by = "Sample") 

ggsave(
  "HDF_Chan_PCA_samples.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Early_senescence"
  )
```

```{r}
DimHeatmap(HDF_Chan_combined, dims=1, cells=500, balanced=TRUE)
DimHeatmap(HDF_Chan_combined, dims=1:15, cells=500, balanced=TRUE)
ElbowPlot(HDF_Chan_combined)

#Use K-nearest neighbours based on PCA distance in FindNeighbours() function based on previously definited dimensions (10) from elbow plot
HDF_Chan_combined <- FindNeighbors(HDF_Chan_combined, dims =1:10)

#group cells using findclusters function
HDF_Chan_combined <- FindClusters(HDF_Chan_combined, resolution = 0.7)

#Look at cluster ID's of first 5 cells
head(Idents(HDF_Chan_combined), 5)
```

## Run UMAPS
```{r}
#Running a UMAP
HDF_Chan_combined <- RunUMAP(HDF_Chan_combined, dims=1:10)

DimPlot(HDF_Chan_combined, reduction = "umap", seed=123, cols =c("#9b5fe0", "#16a5d8", "#60dbe8", "#8bd346", "#efdf48", "#f9a52c", "#d64e12", "#652A0E", "tan3", "peachpuff")) +
    custom_theme(base_size=30)

ggsave(
  "HDF_Chan_UMAP.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Early_senescence"
  )


DimPlot(HDF_Chan_combined, reduction = "umap",group.by = "Sample", seed=123, cols = c( "#24B3A8", "#FF92A5", "#74B72E")) +
  custom_theme(base_size=30)

ggsave(
  "HDF_Chan_UMAP_sample.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Early_senescence"
  )

```
## Testing monocle on the EP/ES/DS data
```{r}
cds <- as.cell_data_set(HDF_Chan_combined)

#Using Seurat's clustering information rather than new Monocle3 clustering
#Assign all the cells to one partition (one supercluster)

recreate.partition <- c(rep(1,length(cds@colData@rownames)))
names(recreate.partition) <- cds@colData@rownames
recreate.partition <- as.factor(recreate.partition)

#Assign this to the partition slot
cds@clusters$UMAP$partitions <- recreate.partition


#Assign the cluster info from Seurat
list_cluster <- HDF_Chan_combined@active.ident
cds@clusters$UMAP$clusters <- list_cluster

cds@int_colData@listData$reducedDims$UMAP <- HDF_Chan_combined@reductions$umap@cell.embeddings


cds <- learn_graph(cds)


get_earliest_principal_node <- function(cds, Sample="EP2D"){
  cell_ids <- which(colData(cds)[, "Sample"] == Sample)
  
  closest_vertex <-
  cds@principal_graph_aux[["UMAP"]]$pr_graph_cell_proj_closest_vertex
  closest_vertex <- as.matrix(closest_vertex[colnames(cds), ])
  root_pr_nodes <-
  igraph::V(principal_graph(cds)[["UMAP"]])$name[as.numeric(names
  (which.max(table(closest_vertex[cell_ids,]))))]
  
  root_pr_nodes
}

cds <- order_cells(cds, root_pr_nodes=get_earliest_principal_node(cds))

plot <- plot_cells(cds,
           color_cells_by = "pseudotime",
           label_cell_groups=FALSE,
           label_leaves=FALSE,
           label_branch_points=FALSE,
           label_groups_by_cluster = FALSE,
           label_roots = FALSE,
           graph_label_size=10,
           group_label_size=10,
           cell_size=1)

plot <- plot + 
  custom_theme(base_size=28)

ggsave("../Output/Early_senescence/pseudotimetrajectory.png", plot, width = 8, height = 7, dpi = 300)


```


# Building an MDA model including the third 'ES' classifier
```{r}
HDF_Chan_combined$new_col <- NULL
#HDF_Chan_combined$New_Col <- NULL
#HDF_Chan_combined$New_Col.1 <- NULL
New_Col <- 1:3235
HDF_Chan_combined@meta.data <- cbind(HDF_Chan_combined@meta.data, New_Col)

HDF_Chan_combined@meta.data$New_Col <- 1:nrow(HDF_Chan_combined@meta.data)
names(HDF_Chan_combined@meta.data)
train <- HDF_Chan_combined@meta.data %>% dplyr::sample_frac(.8)
test  <- dplyr::anti_join(HDF_Chan_combined@meta.data, train, by = 'New_Col')


test <- test$New_Col
train <- train$New_Col

HDF_Chan_combined@meta.data$New_Col[HDF_Chan_combined@meta.data$New_Col %in% test] <- "Test"
HDF_Chan_combined@meta.data$New_Col[HDF_Chan_combined@meta.data$New_Col %in% train] <- "Train"

obj_list <- SplitObject(HDF_Chan_combined, split.by = "New_Col")


#Normalising, scaling and running PCA on training and testing data
training <- obj_list$Train %>% 
  RunUMAP(dims = 1:10)

DimPlot(training, group.by = "Sample", label = FALSE, repel = TRUE, seed=123)

testing <- obj_list$Test%>%
  RunUMAP(dims = 1:10)

DimPlot(testing, group.by = "Sample", label = FALSE, repel=TRUE, seed=123)

##Save training and testing datasets
saveRDS(training, file = "../Output/Seurat_Objects/hdfseuratfiltered_2Dtraining_EPESDS.rds")
saveRDS(testing, file =  "../Output/Seurat_Objects/hdfseuratfiltered_2Dtesting_EPESDS.rds" )

## Note - every time you run the script you will get slightly different results. This is because the training and testing splits are random. If you want to exactly replicate your results, move the training and testing files into 'Input/Seurat_Objects' and load them in each time.

#training <- readRDS("../Input/Seurat_Objects/hdfseuratfiltered_2Dtraining_EPESDS.rds")
#testing <- readRDS("../Input/Seurat_Objects/hdfseuratfiltered_2Dtesting_EPESDS.rds")

```


```{r}
#Using the reference data set to create and evaluate different models - no PC filters
HDF_Chan_combined_scPred <- getFeatureSpace(training, "Sample")


#This model is the mda model
HDF_Chan_combined_scPred <- trainModel(HDF_Chan_combined_scPred, model = "mda")
get_probabilities(HDF_Chan_combined_scPred) %>% head()
get_scpred(HDF_Chan_combined_scPred)
HDF_Chan_combined_scPred
```

## Model testing
```{r}
#MDA testing 
query <- scPredict(testing, HDF_Chan_combined_scPred, recompute_alignment = FALSE)

#Plotting UMAP of predicted senescent cells
query <- RunUMAP(query, reduction = "scpred", dims = 1:10)
DimPlot(query, group.by = "scpred_prediction", label = FALSE, repel = TRUE, seed = 123, cols = c("darkblue", "deeppink3", "#00CA00", "antiquewhite4"))

ggsave(
  "UMAP_scpred_MDA_ES.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Early_senescence"
  )

DimPlot(query, group.by = "Sample", label = FALSE, repel = TRUE, seed = 123)
ggsave(
  "UMAP_scpred_MDA_ES_SAMPLE.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Early_senescence/2D_SVM"
  )
```

## Confusion matrix plot
```{r}
query_df <- query@meta.data
query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="DS2D" & query_df$scpred_prediction=="DS2D" ~ "DS2DT"))

ds2dt <- nrow(query_df[!is.na(query_df$Status),])

query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="DS2D" & query_df$scpred_prediction=="EP2D" ~ "EP2DF"))

ep2df <- nrow(query_df[!is.na(query_df$Status),])

query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="DS2D" & query_df$scpred_prediction=="PDL50" ~ "PDL50F"))

pdl50f <- nrow(query_df[!is.na(query_df$Status),])

query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="EP2D" & query_df$scpred_prediction=="EP2D" ~ "EP2DT"))

ep2dt <- nrow(query_df[!is.na(query_df$Status),])

query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="EP2D" & query_df$scpred_prediction=="PDL50" ~ "PDL50EPF"))

PDL50EPF <- nrow(query_df[!is.na(query_df$Status),])


query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="PDL50" & query_df$scpred_prediction=="PDL50" ~ "PDL50T"))

PDL50T <- nrow(query_df[!is.na(query_df$Status),])


query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="EP2D" & query_df$scpred_prediction=="DS2D" ~ "DS2DF"))

ds2df <- nrow(query_df[!is.na(query_df$Status),])

query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="PDL50" & query_df$scpred_prediction=="DS2D" ~ "DS2DESF"))

ds2desf <- nrow(query_df[!is.na(query_df$Status),])

query_df <- query_df %>%
  mutate(Status = case_when(query_df$Sample=="PDL50" & query_df$scpred_prediction=="EP2D" ~ "EP2DESF"))

ep2desf <- nrow(query_df[!is.na(query_df$Status),])

unassigned <- length(which(query_df$scpred_prediction == "unassigned"))

#Plotting into confusion matrix
table <- matrix(c(ds2dt, ep2df, pdl50f, ds2df, ep2dt, PDL50EPF, ds2desf, ep2desf, PDL50T),ncol=3,byrow=TRUE)
colnames(table) <- c("DS", "EP", "PDL50")
rownames(table) <- c("DS", "EP", "PDL50")
table <- as.table(table)
table


perc <- prop.table(table, 1)*100
perc

perc <- round(perc, digits=2)

cm <- confusionMatrix(perc)

# Drawing confusion matrix into visual table

draw_confusion_matrix <- function(cm, output_file) {

  draw_matrix_plot <- function() {
  layout(matrix(c(1,1)))
  par(mar=c(2,2,2,2))
  plot(c(100, 345), c(200, 470), type = "n", xlab="", ylab="", xaxt='n', yaxt='n')
  title('MDA Confusion Matrix (%)', cex.main=2)

  # create the matrix 
  rect(150, 420, 200, 360, col='#32CD32')
  text(175, 440, 'DS', cex=2)
  rect(205, 420, 255, 360, col='#FF4500')
  text(230, 440, 'EP', cex=2)
  rect(260, 420, 310, 360, col='#FF4500')
  text(285, 440, 'ES', cex=2)
  
  text(125, 325, 'Predicted', cex=2, srt=90, font=2)
  text(230, 465, 'Actual', cex=2, font=2)
  rect(150, 290, 200, 350, col='#FF4500')
  rect(205, 290, 255, 350, col='#32CD32')
  text(140, 390, 'DS', cex=2, srt=90)
  text(140, 320, 'EP', cex=2, srt=90)
  rect(260, 290, 310, 350, col='#FF4500')
  rect(150, 220, 200, 280, col='#FF4500')
  rect(205, 220, 255, 280, col='#FF4500')
  rect(260, 220, 310, 280, col='#32CD32')
  text(140, 250, 'ES', cex=2, srt=90)
  
  
  

  # add in the cm results 
  res <- as.numeric(cm$table)
  text(176, 393, res[1], cex=2, font=2, col='white')
  text(176, 323, res[2], cex=2, font=2, col='white')
  text(176, 253, res[3], cex=2, font=2, col='white')
  text(230, 393, res[4], cex=2, font=2, col='white')
  text(230, 323, res[5], cex=2, font=2, col='white')
  text(230, 253, res[6], cex=2, font=2, col='white')
  text(284, 393, res[7], cex=2, font=2, col='white')
  text(284, 323, res[8], cex=2, font=2, col='white')
  text(284, 253, res[9], cex=2, font=2, col='white')
  text(120, 200, 'Unassigned =', cex=1, font=1, col='black')
  text(155, 201, unassigned, cex=1, font=1, col='black')
}
  # Call the function to draw the matrix plot
  draw_matrix_plot()

  # Save the plot as a PNG file
  png(filename = output_file, width = 800, height = 600)  # Set appropriate width and height
  draw_matrix_plot()
  dev.off()  # Turn off the PNG device
}


your_confusion_matrix <- cm
output_file_path <- "../Output/Early_senescence/CM2D_MDA_testing.png"
draw_confusion_matrix(your_confusion_matrix, output_file_path)

```

# Running ES model on whole Chan dataset

```{r}
query <- NormalizeData(Chan_seurat2)
query <- scPredict(query, HDF_Chan_combined_scPred)

#Plotting UMAP of predicted senescent cells
DimPlot(query, group.by = "scpred_prediction", label = FALSE, repel = TRUE, seed = 123, cols = c("darkblue", "deeppink3", "#228c22", "antiquewhite4")) +
    custom_theme(base_size=20)
  

ggsave(
  "UMAP_Chan_scpred_ESmodel.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Early_senescence"
  )
```

## Investigating the percentages of predicted early senescent cells from each group
```{r}
query2 <- query@meta.data
table <- table(query2[, c("PDL", "scpred_prediction")])
table
mat <- as.data.frame.matrix(table)
mat

percentages <- (mat/rowSums(mat))*100
percentages
percentages <- na.omit(percentages)
percentages
as.matrix(percentages)
percentages <- percentages[ order(row.names(percentages)), ]
v <- rownames(percentages)

barplot_data <- percentages$PDL50
barplot_names <- v

png(filename = "../Output/Early_Senescence/scPred2D_Chan_barplot_ESperPDL.png", width = 800, height = 600)  # Set appropriate width and height
barplot(barplot_data, names = barplot_names,cex.names = 1.8, cex.axis = 1.7, cex.lab = 1.7, cex.main=2.5, col = "darkblue", main = "Percentage of predicted senescent cells",
        xlab = "Population Doublings", ylab = "Percentage", ylim = c(0, 100))
dev.off() 

```

## Investigating new ES model on EP DS cells 
```{r}
query <- NormalizeData(hdfseurat_filt_2D)
query <- scPredict(query, HDF_Chan_combined_scPred)

#Plotting UMAP of predicted senescent cells
DimPlot(query, group.by = "scpred_prediction", label = FALSE, repel = TRUE, seed = 123, cols = c("darkblue", "deeppink3", "#00CA00", "antiquewhite4"))
```

## Investigating the percentages of predicted early senescent cells from each group
```{r}
query2 <- query@meta.data
table <- table(query2[, c("Sample", "scpred_prediction")])
table
mat <- as.data.frame.matrix(table)
mat

percentages <- (mat/rowSums(mat))*100
percentages
percentages <- na.omit(percentages)
percentages
as.matrix(percentages)
percentages <- percentages[ order(row.names(percentages)), ]
v <- rownames(percentages)

barplot_data <- percentages$PDL50
barplot_names <- v

png(filename = "../Output/Early_Senescence/scPred2D_Chan_barplot_ESperEPDS.png", width = 800, height = 600)  # Set appropriate width and height
barplot(barplot_data, names = barplot_names, cex.names = 1.8, cex.axis = 1.7, cex.lab = 1.7, cex.main=2.5, col = "darkblue", main = "Percentage of predicted early senescent cells",
         ylab = "Percentage", ylim = c(0, 100))
dev.off() 


```
### Session information
```{r session_info_99}
sessionInfo()
```

This document was processed on: `r Sys.Date()`.