---
title: "Tabib"
author: "BH"
date: '2022-04-25'
output: html_document
---
```{r}
Tabib <- readRDS("../Inputfiles/Tabib_SeuratObject_AD23-Clusters.rds")
Idents(object = Tabib) <- Tabib$Tabib_name

#Extracting just the fibroblasts
Tabib_fibs <- subset(x = Tabib, idents = c("0_Fibroblast", "3_Fibroblast", "4_Fibroblast"))
```
## Applying same QC metrics as my data to test for outliers. 
nCount_RNA > 1000
nFeature_RNA > 400
percent.mt <= 12


```{r}
#Adding mitochondrial percentage and number of genes per UMI
Tabib_fibs$log10GenesPerUMI <- log10(Tabib_fibs$nFeature_RNA) / log10(Tabib_fibs$nCount_RNA)

Tabib_fibs[["percent.mt"]] <- PercentageFeatureSet(Tabib_fibs, pattern = "^MT-")

#Visualise the number of cells per sample
Tabib_fibs@meta.data %>% 
  	ggplot(aes(x=Tabib_name, fill=Tabib_name)) + 
  	geom_bar() +
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("NCells")

#Save plot
ggsave(
  "NumbersPerSample.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )

```

## Visualising the number of UMIs per cell (nCount_RNA)

```{r}
#Number of UMIs (transcripts) per cell
Tabib_fibs@meta.data %>% 
  	ggplot(aes(color=Tabib_name, x=nCount_RNA, fill= Tabib_name)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	ylab("Cell density") +
  	geom_vline(xintercept = 1000)

#Save plot
ggsave(
  "UMIspercell.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )


```

## Visualising number of genes per cell (nFeature_RNA)
```{r}
# Visualize the distribution of genes detected per cell via histogram
Tabib_fibs@meta.data %>% 
  	ggplot(aes(color=Tabib_name, x=nFeature_RNA, fill= Tabib_name)) + 
  	geom_density(alpha = 0.2) + 
  	theme_classic() +
  	scale_x_log10() + 
  	geom_vline(xintercept = c(400))

#Save plot
ggsave(
  "hist_featureRNA.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )

# Visualize the distribution of genes detected per cell via boxplot
Tabib_fibs@meta.data %>% 
  	ggplot(aes(x=Tabib_name, y=log10(nFeature_RNA), fill=Tabib_name)) + 
  	geom_boxplot() + 
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("NCells vs NGenes")

#Save plot
ggsave(
  "boxplot_featureRNA.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )
```

## Combining number of genes, UMIs and mitochondrial percentage, to show thresholding boundaries
```{r}
Tabib_fibs@meta.data %>% 
  	ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) + 
  	geom_point() + 
	scale_colour_gradient(low = "gray90", high = "black") +
  	stat_smooth(method=lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = 1000) +
  	geom_hline(yintercept = 400) +
  	facet_wrap(~Tabib_name)

#Save plot
ggsave(
  "threshold_feature_count_mito.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )
```

## Mitochondrial gene expression per cell

```{r}
# Visualize the distribution of mitochondrial gene expression detected per cell
Tabib_fibs@meta.data %>% 
  	ggplot(aes(color=Tabib_name, x=percent.mt, fill=Tabib_name)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = 12)

#Save plot
ggsave(
  "mitochondrial_hist.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )
```

## Visualising the proportion of genes per UMI
```{r}
#Genes per UMI
Tabib_fibs@meta.data %>%
  	ggplot(aes(x=log10GenesPerUMI, color = Tabib_name, fill=Tabib_name)) +
  	geom_density(alpha = 0.2) +
  	theme_classic() +
  	geom_vline(xintercept = 0.8)
```

#Subsetting seurat object
```{r}
Tabib_fibs2 <- subset(x=Tabib_fibs, 
                           (percent.mt <=12) &
                            (nFeature_RNA >=400) &
                            (nCount_RNA >= 1000))

before <- table(Tabib_fibs@meta.data['Tabib_name'])
after <- table(Tabib_fibs2@meta.data['Tabib_name'])

before <- as.data.frame(before)
after <- as.data.frame(after)

combo <- merge(before, after, by = "Tabib_name")


combo2 <- melt(combo)



ggplot(combo2, aes(fill=variable, y=value, x=Tabib_name)) + 
    geom_bar(position="dodge", stat="identity")

#Save plot
ggsave(
  "beforeandafterfilt.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )

```

## Finding top variable genes within the dataset

```{r}
#Selecting highly variable features - high cell to cell variation
Tabib_fibs2 <- FindVariableFeatures(Tabib_fibs2, selection.method ="vst", nfeatures=2000)

#Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(Tabib_fibs2), 10)
top10

top2000 <- head(VariableFeatures(Tabib_fibs2), 2000)
top2000

#Plot variable features without labels
plot1 <- VariableFeaturePlot(Tabib_fibs2)

#Plot variable features with labels
plot2 <- LabelPoints(plot=plot1, points = top10, repel=TRUE)
plot1
plot2

ggsave(
  "topvariablefeatures.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )
```

## Scaling the data before dimensionality reduction. Each gene is centred to a mean of 0, and scaled by the standard deviation of each gene.
```{r}
all.genes <- rownames(Tabib_fibs2)
Tabib_fibs2 <- ScaleData(Tabib_fibs2, features = all.genes)

#Performing a PCA
Tabib_fibs2 <- RunPCA(Tabib_fibs2, features=VariableFeatures(object = Tabib_fibs2))

#Plotting the PCA
PCA_no_label <- DimPlot(Tabib_fibs2, reduction = "pca", seed = 123, combine=TRUE) + NoLegend()
PCA_Sample_label <- DimPlot(Tabib_fibs2, reduction = "pca", seed = 123, group.by = "Tabib_name")

PCA_no_label
ggsave(
  "PCA_nolabel.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )

PCA_Sample_label
ggsave(
  "PCA_PDLlabel.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )


```

## Heatmaps to investigate the genes influencing the top PCs
```{r}
DimHeatmap(Tabib_fibs2, dims=1, cells=500, balanced=TRUE)
DimHeatmap(Tabib_fibs2, dims=1:15, cells=500, balanced=TRUE)
ElbowPlot(Tabib_fibs2)
```

## Clustering the cells using KNN
```{r}
#Use K-nearest neighbours based on PCA distance in FindNeighbours() function based on previously definited dimensions (10) from elbow plot
Tabib_fibs2 <- FindNeighbors(Tabib_fibs2, dims =1:10)

#group cells using findclusters function
Tabib_fibs2 <- FindClusters(Tabib_fibs2, resolution = 0.2)

#Look at cluster ID's of first 5 cells
head(Idents(Tabib_fibs2), 5)
```

## Plot UMAP
```{r}
#Running a UMAP
Tabib_fibs2 <- RunUMAP(Tabib_fibs2, dims=1:10)

DimPlot(Tabib_fibs2, reduction = "umap", seed = 123, cols = c("#9b5fe0", "#16a5d8", "#60dbe8", "#8bd346", "#efdf48")) +
  custom_theme(base_size=20)
ggsave(
  "UMAP_clusters_all.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )

DimPlot(Tabib_fibs2, reduction = "umap", seed = 123, group.by="Tabib_name", cols = c("#2e5266", "#ecb0e1", "#832232", "#a6a867", "#bc5f04", "#f9a52c")) + 
  custom_theme(base_size=20)
ggsave(
  "UMAP_clusters_all_Sample.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )
```

##Testing ScPred 2D ML models on Chan datasets
```{r}
#MDA radial model
query <- NormalizeData(Tabib_fibs2)
query <- scPredict(query, hdf_mda2D)

DimPlot(query, group.by = "Tabib_name", label = FALSE, repel = TRUE, label.size=2, seed=123) + 
  custom_theme(base_size=20)
        
       
ggsave(
  "UMAP_PDL_notert_Sample.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )


DimPlot(query, group.by = "scpred_prediction", label = FALSE, repel = TRUE, seed=123, cols = c("darkblue", "deeppink3", "antiquewhite4")) + 
  custom_theme(base_size=20)

ggsave(
  "UMAP_scpred_EPDS_2D.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )
```
## Testing EP/ES/DS model
```{r}
#MDA radial model
query <- NormalizeData(Tabib_fibs2)
query <- scPredict(query, HDF_Chan_combined_scPred)

DimPlot(query, group.by = "Tabib_name", label = FALSE, repel = TRUE, label.size=2, seed=123) + 
  custom_theme(base_size=20)
      
DimPlot(query, group.by = "scpred_prediction", label = FALSE, repel = TRUE, seed=123, cols = c("darkblue", "deeppink3", "#228c22", "antiquewhite4")) + 
  custom_theme(base_size=20)

ggsave(
  "UMAP_scpred_EPESDS_2D.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )

#Colour UMAP by age
DimPlot(query, group.by = "age", label = FALSE, repel = TRUE, seed=123, cols = c("#cc889d", "#f5c5a0", "#7dae9b", "#8abfe8", "#5c7998", "#7b78b8")) + 
  custom_theme(base_size=20)

ggsave(
  "UMAP_scpred_EPESDS_2D_AGE.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )
```

## Investigating the percentage of DS predicted cells per age

```{r}
query2 <- query@meta.data
table <- table(query2[, c("age", "scpred_prediction")])
table
mat <- as.data.frame.matrix(table)
mat

percentages <- (mat/rowSums(mat))*100
percentages
percentages <- na.omit(percentages)
percentages
as.matrix(percentages)
percentages <- percentages[ order(row.names(percentages)), ]
v <- rownames(percentages)


barplot_data <- percentages$DS2D
barplot_names <- v

png(filename = "../Output/Tabib/scPred2D_Tabib_barplot_percentageDS.png", width = 800, height = 600)  # Set appropriate width and height
barplot(barplot_data, names = barplot_names, cex.names = 1.8, cex.axis = 1.7, cex.lab = 1.7, cex.main=2.5, col = "darkblue", main = "Percentage of predicted deeply senescent cells", xlab="Population Doublings",
         ylab = "Percentage", ylim = c(0, 100))
dev.off() 

```
## Investigating the percentage of DS predicted cells per age

```{r}
query2 <- query@meta.data
table <- table(query2[, c("age", "scpred_prediction")])
table
mat <- as.data.frame.matrix(table)
mat

percentages <- (mat/rowSums(mat))*100
percentages
percentages <- na.omit(percentages)
percentages
as.matrix(percentages)
percentages <- percentages[ order(row.names(percentages)), ]


df <- tibble::rownames_to_column(percentages, "age")
perc <- melt(df)
v <- perc$age

#convert 'variable' to factor and specify level order
perc$variable <- factor(perc$variable, levels=c('unassigned', 'EP2D', 'PDL50', 'DS2D'))

#create stacked bar chart
ggplot(perc, aes(x=age, y=value, fill=variable)) + 
    geom_bar(position='stack', stat='identity') +
  scale_fill_manual(values=c("antiquewhite4", "deeppink3", "#228c22", "darkblue")) +
  custom_theme(base_size=20)

ggsave("barplot_stacked_2Dpred_age.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Tabib"
  )

```

### Session information
```{r session_info_99}
sessionInfo()
```

This document was processed on: `r Sys.Date()`.
