---
title: "7_Teo"
author: "BH"
date: "2024-03-15"
output: html_document
---
## Loading barcodes and conditions, and counts matrix for publicly available paper: Teo, Y.V., Rattanavirotkul, N., Olova, N., Salzano, A., Quintanilla, A., Tarrats, N., Kiourtis, C., Müller, M., Green, A.R., Adams, P.D. and Acosta, J.C., 2019. Notch signaling mediates secondary senescence. Cell reports, 27(4), pp.997-1007.

This will allow the model to be tested on an alternative trigger: Oncogene Induced Senescence (OIS/RIS), and paracrine senescence (GFP).

```{r}
#install.packages("GEOquery")
library(GEOquery)
metadata <- read.csv("../Inputfiles/barcodes_Teo.csv")
counts <- read.table("../Inputfiles/GSE115301_Growing_Sen_10x_count.txt")

rownames(metadata) <- metadata$X
metadata = subset(metadata, select = -c(1))

#Creating Seurat object
teo <- CreateSeuratObject(counts = counts)   
teo_seurat <- AddMetaData(teo, metadata)

teo_seurat@meta.data <- merge(teo_seurat@meta.data, metadata,
                          by = 'row.names', all = TRUE)

teo_seurat@meta.data = subset (teo_seurat@meta.data, select = -c(6))
rownames(teo_seurat@meta.data) <- teo_seurat@meta.data$Row.names
teo_seurat@meta.data = subset (teo_seurat@meta.data, select = -c(1))

saveRDS(teo_seurat, file = "../Output/Seurat_Objects/Teo_seurat.RDS")
```


```{r}
Teo_seurat <- readRDS(file = "../Output/Seurat_Objects/Teo_seurat.RDS")

#Generating log10 genes per UMI for QC
Teo_seurat$log10GenesPerUMI <- log10(Teo_seurat$nFeature_RNA) / log10(Teo_seurat$nCount_RNA)

#Generating percentage mitochondrial features for QC
Teo_seurat[["percent.mt"]] <- PercentageFeatureSet(Teo_seurat, pattern = "^MT-")


#Visualise the number of cells per sample
Teo_numbers <- Teo_seurat@meta.data %>% 
  	ggplot(aes(x = Condition2.x, fill = Condition2.x)) + 
  	geom_bar() + 
  scale_fill_manual(values = c("#C22A3D", "darkgreen", "#2232A1"))+
  	theme_classic() +
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("NCells") + custom_theme(base_size=20) +
  labs(x = "Condition", fill="Condition")


Teo_numbers

ggsave(
  "TeoNumbersPerSample.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Teo_QC"
  )
```

```{r}
#Number of UMIs (transcripts) per cell
Teo_UMI <- Teo_seurat@meta.data %>% 
  	ggplot(aes(x=nCount_RNA, fill= Condition2.x)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	ylab("Cell density") +
  	geom_vline(xintercept = 1000)+
  labs(fill="Condition")+
  custom_theme(base_size=20)

Teo_UMI

ggsave(
  "TeoNumberofUMI.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Teo_QC"
  )
```


```{r}
# Visualize the distribution of genes detected per cell via histogram
Teo_features <- Teo_seurat@meta.data %>% 
  	ggplot(aes(x=nFeature_RNA, fill= Condition2.x)) + 
  	geom_density(alpha = 0.2) + 
  	theme_classic() +
  	scale_x_log10() + 
  	geom_vline(xintercept = 400)+
  labs(fill="Condition")+
  custom_theme(base_size=20)

Teo_features

ggsave(
  "Teofeaturespercell.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Teo_QC"
  )
```

```{r}
# Visualize the distribution of genes detected per cell via boxplot
Boxplot_genespercell_Teo <- Teo_seurat@meta.data %>% 
  	ggplot(aes(x=Condition2.x, y=log10(nFeature_RNA), fill=Condition2.x)) + 
  	geom_boxplot() + 
  	theme_classic() +
    scale_fill_manual(values = c("#C22A3D", "darkgreen", "#2232A1"))+
  	theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  	theme(plot.title = element_text(hjust=0.5, face="bold")) +
  	ggtitle("NCells vs NGenes")+
  custom_theme(base_size=20)+
  labs(fill="Condition", x="Condition")

Boxplot_genespercell_Teo

ggsave(
  "Teofeaturespercell_boxplot.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Teo_QC"
  )
```

```{r}
Feature_count_mt <- Teo_seurat@meta.data %>% 
  	ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) + 
  	geom_point() + 
	scale_colour_gradient(low = "gray90", high = "black") +
  	stat_smooth(method=lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = 1000) +
  	geom_hline(yintercept = 400) +
  	facet_wrap(~Condition2.x)+
  custom_theme(base_size=10)

Feature_count_mt

ggsave(
  "Teofeature_count_mt.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Teo_QC"
  )
```


```{r}
# Visualize the distribution of mitochondrial gene expression detected per cell
percentmt_Teo<- Teo_seurat@meta.data %>% 
  	ggplot(aes(x=percent.mt, fill=Condition2.x)) + 
  	geom_density(alpha = 0.2) + 
  	scale_x_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = 12)+
  custom_theme(base_size=20)+
  labs(fill="Condition")

percentmt_Teo

ggsave(
  "Teo_mt.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Teo_QC"
  )
```


```{r}
#Log10 genes per UMI
teo_log10umi <- Teo_seurat@meta.data %>%
  	ggplot(aes(x=log10GenesPerUMI, fill=Condition2.x)) +
  	geom_density(alpha = 0.2) +
  	theme_classic() +
  	geom_vline(xintercept = 0.8)+
  custom_theme(base_size=20)+
  labs(fill="Condition")

teo_log10umi

ggsave(
  "Teo_log10umi.tiff",
  plot = last_plot(),
  device = "tiff",
  path = "../Output/Teo_QC"
  )

```
Now that we have visualized the various metrics, we can decide on the thresholds to apply which will result in the removal of low quality cells. 

To remove potentially apoptotic or 'low quality' cells we discarded cells with less than 400 expressed genes, and/or more than 12% mitochondrial reads. We also filtered out cells with a UMI count less than 1000. 

nCount_RNA > 1000
nFeature_RNA > 400
percent.mt <= 12
```{r}
Teo_seurat2 <- subset(x=Teo_seurat, 
                           subset= (percent.mt <=12) &
                            (nFeature_RNA >=400) &
                            (nCount_RNA >= 1000))


df.list <- list(Original=Teo_seurat@meta.data, Filtered=Teo_seurat2@meta.data)
dat <- stack(lapply(df.list, `[[`, "Condition2.x"))
ggplot(dat, aes(x=values, fill=ind)) + geom_bar(position="dodge")

#LogNormalize data
Teo_seurat2 <- NormalizeData(Teo_seurat2)
```

```{r}
#Selecting highly variable features - high cell to cell variation
Teo_seurat2 <- FindVariableFeatures(Teo_seurat2, selection.method ="vst", nfeatures=2000)

#Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(Teo_seurat2), 10)
top10

top2000 <- head(VariableFeatures(Teo_seurat2), 2000)
top2000

#Plot variable features without labels
plot1 <- VariableFeaturePlot(Teo_seurat2)
#Plot variable features with labels
plot2 <- LabelPoints(plot=plot1, points = top10, repel=TRUE)
plot1
plot2

```


```{r}
all.genes <- rownames(Teo_seurat2)
Teo_seurat2 <- ScaleData(Teo_seurat2, features = all.genes)

#Performing a PCA
Teo_seurat2 <- RunPCA(Teo_seurat2, features=VariableFeatures(object = Teo_seurat2))

DimPlot(Teo_seurat2, reduction = "pca")
tiff("../Output/Teo/PCA.tiff")
dev.off()
```
```{r}
DimHeatmap(Teo_seurat2, dims=1, cells=500, balanced=TRUE)
DimHeatmap(Teo_seurat2, dims=1:15, cells=500, balanced=TRUE)
ElbowPlot(Teo_seurat2)

#Use K-nearest neighbours based on PCA distance in FindNeighbours() function based on previously definited dimensions (10) from elbow plot
Teo_seurat2 <- FindNeighbors(Teo_seurat2, dims =1:10)

#group cells using findclusters function
Teo_seurat2 <- FindClusters(Teo_seurat2, resolution = 0.5)

#Look at cluster ID's of first 5 cells
head(Idents(Teo_seurat2), 5)
```
```{r}
#Running a UMAP
Teo_seurat2 <- RunUMAP(Teo_seurat2, dims=1:10)

DimPlot(Teo_seurat2, reduction = "umap")
tiff("../Output/Teo/UMAP_clusters.tiff",width=3300,height=2400,res=400)


DimPlot(Teo_seurat2, reduction = "umap",group.by = "Condition2.x")
tiff("../Output/Teo/UMAP_samples.tiff",width=3300,height=2400,res=400)
```

##Running SenPred 2D EP/ES/DS on Teo data
```{r}
query <- NormalizeData(Teo_seurat2)
query <- scPredict(query, HDF_Chan_combined_scPred)


reference <- Teo_seurat2 %>% 
  NormalizeData() %>% 
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA() %>% 
  RunUMAP(dims = 1:30)

DimPlot(query, group.by = "Condition2.x", label = FALSE, repel = TRUE, label.size=2)
DimPlot(query, group.by = "scpred_prediction", label = FALSE, repel = TRUE)

query <- RunUMAP(query, reduction = "scpred", dims = 1:30)
DimPlot(query, group.by = "scpred_prediction", label = TRUE, repel = TRUE,  label.size = 4)
DimPlot(query, group.by = "Condition2.x", label = TRUE, repel = TRUE,  label.size = 2)
FeaturePlot(query, c("scpred_DS3D", "scpred_EP3D", "unassigned"))



```

```{r}
# Subset the Seurat object based on scpred_prediction
filtered_seurat <- subset(Teo_seurat2, subset = Condition2.x == "GFP")

# Extract gene counts for filtered cells
gene_counts <- filtered_seurat@assays$RNA@counts

# Define gene list - TNF is missing
gene_list <- c("CDKN1A", "CDKN2A", "IGFBP3", "SERPINE1", "CXCL12")

# Check if gene names in gene_list are present in row names (genes)
missing_genes <- gene_list[!gene_list %in% rownames(gene_counts)]
if (length(missing_genes) > 0) {
  stop(paste("Genes missing in gene counts:", paste(missing_genes, collapse = ", ")))
}

# Filter gene counts based on the gene list
gene_counts <- gene_counts[gene_list, ]
gene_counts <- as.matrix(gene_counts)
# Transpose gene counts matrix
gene_counts_transposed <- t(gene_counts)
gene_counts_transposed <- as.matrix(gene_counts_transposed)

# Calculate correlation matrix
correlation_matrix <- cor(gene_counts_transposed)

# Print correlation matrix
print(correlation_matrix)

# Define breaks and color palette
breaks <- unique(c(seq(-1,-0.0001,length=50),seq(-0.00001,0.01,length=50), seq(0.011,1,length=50)))
my_palette <- colorRampPalette(c("blue","white", "white", "red"))(length(breaks)-1)

library("gplots")
# Assume combined3 is the correlation matrix
heatmap <- heatmap.2(correlation_matrix, 
                     col = my_palette, 
                     breaks = breaks, 
                     key = TRUE, 
                     trace = "none", 
                     main = "Marker Correlation_Teo_RIS", 
                     key.title = "Correlation",   
                     cexRow = 0.6, 
                     cexCol = 0.6,    
                     distfun = function(x) dist(x, method = "euclidean"),
                     hclust = function(x) hclust(x, method = "ward.D2"), 
                     symkey = FALSE, 
                     Rowv = FALSE,
                     Colv = FALSE,
                     density.info = c("none"))


```

```{r}
query <- NormalizeData(Teo_seurat2)
query <- scPredict(query, HDF_Chan_combined_scPred)


reference <- Teo_seurat2 %>% 
  NormalizeData() %>% 
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA() %>% 
  RunUMAP(dims = 1:30)

DimPlot(query, group.by = "Condition2.x", label = FALSE, repel = TRUE, label.size=2)
DimPlot(query, group.by = "scpred_prediction", label = FALSE, repel = TRUE)

query <- RunUMAP(query, reduction = "scpred", dims = 1:30)
DimPlot(query, group.by = "scpred_prediction", label = TRUE, repel = TRUE,  label.size = 4)
DimPlot(query, group.by = "Condition2.x", label = TRUE, repel = TRUE,  label.size = 2)
FeaturePlot(query, c("scpred_DS3D", "scpred_EP3D", "unassigned"))



```


```{r}
query2 <- query@meta.data
table <- table(query2[, c("Condition2.x", "scpred_prediction")])
table
mat <- as.data.frame.matrix(table)
mat

percentages <- (mat/rowSums(mat))*100
percentages
percentages <- na.omit(percentages)
percentages
as.matrix(percentages)
percentages <- percentages[ order(row.names(percentages)), ]
v <- rownames(percentages)

percentages
df <- tibble::rownames_to_column(percentages, "Condition")
percentage2 <- melt(df)

ggplot(percentage2, aes(x=Condition, y=value, fill=variable)) + 
    geom_bar(position='stack', stat='identity') +
  scale_fill_manual(values=c("antiquewhite4", "deeppink3", "#228c22", "darkblue" )) 

```
#Building machine learning models with Teo growing and RIS
```{r}
Teo_RIS_grow <- subset(Teo_seurat, Condition2.x == "RIS" | Condition2.x == "Growing")

Teo_RIS_grow_filt <- subset(x=Teo_RIS_grow, 
                            (nFeature_RNA >=500) &
                            (nCount_RNA >= 1000))

#Seperating into training and testing sets
new_col <- 1:447

Teo_RIS_grow_filt@meta.data <- cbind(Teo_RIS_grow_filt@meta.data, new_col)

Teo_RIS_grow_filt@meta.data$new_col <- 1:nrow(Teo_RIS_grow_filt@meta.data)
train <- Teo_RIS_grow_filt@meta.data %>% dplyr::sample_frac(.8)
test  <- dplyr::anti_join(Teo_RIS_grow_filt@meta.data, train, by = 'new_col')

test <- test$new_col
train <- train$new_col

Teo_RIS_grow_filt@meta.data$new_col[Teo_RIS_grow_filt@meta.data$new_col %in% test] <- "Test"
Teo_RIS_grow_filt@meta.data$new_col[Teo_RIS_grow_filt@meta.data$new_col %in% train] <- "Train"

obj_list <- SplitObject(Teo_RIS_grow_filt, split.by = "new_col")

training <- obj_list$Train %>% 
  NormalizeData() %>% 
   FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA() %>% 
  RunUMAP(dims = 1:10)

DimPlot(training, group.by = "Condition2.x", label = FALSE, repel = TRUE)

testing <- obj_list$Test%>%
  NormalizeData() %>%
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA() %>% 
  RunUMAP(dims = 1:10)

DimPlot(testing, group.by = "Condition2.x", label = FALSE, repel=TRUE)


```
```{r}
#Using the training data set to create and evaluate different models - no PC filters
Teo_scpred <- getFeatureSpace(training, "Condition2.x")


#This model is the svm Radial model
Teo_trained <- trainModel(Teo_scpred)
get_probabilities(Teo_trained) %>% head()
Teo_svmradial <- get_scpred(Teo_trained)
Teo_svmradial

```
```{r}
query <- NormalizeData(testing)
query <- scPredict(query, Teo_svmradial, recompute_alignment = FALSE, threshold = 0.55)

```

```{r}
query_df <- query@meta.data
query_df <- query_df %>%
  mutate(Status = case_when(query_df$Condition2.x=="RIS" & query_df$scpred_prediction=="RIS" ~ "RIST"))

ds2dt <- nrow(query_df[!is.na(query_df$Status),])

query_df <- query_df %>%
  mutate(Status = case_when(query_df$Condition2.x=="RIS" & query_df$scpred_prediction=="Growing" ~ "GrowingF"))

ep2df <- nrow(query_df[!is.na(query_df$Status),])

query_df <- query_df %>%
  mutate(Status = case_when(query_df$Condition2.x=="Growing" & query_df$scpred_prediction=="Growing" ~ "GrowingT"))

ep2dt <- nrow(query_df[!is.na(query_df$Status),])


query_df <- query_df %>%
  mutate(Status = case_when(query_df$Condition2.x=="Growing" & query_df$scpred_prediction=="RIS" ~ "RISF"))

ds2df <- nrow(query_df[!is.na(query_df$Status),])

#Plotting into confusion matrix

table <- matrix(c(ds2dt, ep2df, ds2df, ep2dt),ncol=2,byrow=TRUE)
colnames(table) <- c("DS", "EP")
rownames(table) <- c("DS", "EP")
table <- as.table(table)
table

```

```{r}
#Can OIS predict 2D data?
hdfdeb_filt_2D <- subset(hdfbeth_filt, Sample == "DS2D" | Sample == "EP2D")
query2D <- NormalizeData(hdfdeb_filt_2D)
query2D <- scPredict(query2D, Teo_svmradial)

DimHeatmap(query2D, dims=1, cells=500, balanced=TRUE)
DimHeatmap(query2D, dims=1:15, cells=500, balanced=TRUE)
ElbowPlot(query2D)

#Use K-nearest neighbours based on PCA distance in FindNeighbours() function based on previously definited dimensions (10) from elbow plot
query2D <- FindNeighbors(query2D, dims =1:10)

#group cells using findclusters function
query2D <- FindClusters(query2D, resolution = 0.5)
```

```{r}
#Running a UMAP
hdfdeb_filt_2D <- RunUMAP(query2D, dims=1:10)

DimPlot(hdfdeb_filt_2D, reduction = "umap")
tiff("../Results/UMAP_clusters_2D.tiff",width=3300,height=2400,res=400)
DimPlot(hdfdeb_filt_2D, reduction = "umap")


DimPlot(hdfdeb_filt_2D, reduction = "umap",group.by = "Sample")
tiff("../Results/UMAP_samples_2D.tiff",width=3300,height=2400,res=400)
DimPlot(hdfdeb_filt_2D, reduction = "umap",group.by = "Sample")

DimPlot(query2D, reduction = "umap",group.by = "scpred_prediction")
tiff("../Results/UMAP_phase_2D.tiff",width=3300,height=2400,res=400)
DimPlot(hdfdeb_filt_2D, reduction = "umap",group.by = "Phase")

DimPlot(hdfdeb_filt_2D, reduction = "umap", group.by = "scpred_prediction")
```

```{r}
query_df2D <- query2D@meta.data

query_df2D <- query_df2D %>%
  mutate(Status = case_when(query_df2D$Sample=="DS2D" & query_df2D$scpred_prediction=="RIS" ~ "RIST"))

ds2dt <- nrow(query_df2D[!is.na(query_df2D$Status),])

query_df2D <- query_df2D %>%
  mutate(Status = case_when(query_df2D$Sample=="DS2D" & query_df2D$scpred_prediction=="Growing" ~ "GrowingF"))

ep2df <- nrow(query_df2D[!is.na(query_df2D$Status),])

query_df2D <- query_df2D %>%
  mutate(Status = case_when(query_df2D$Sample=="EP2D" & query_df2D$scpred_prediction=="Growing" ~ "GrowingT"))

ep2dt <- nrow(query_df2D[!is.na(query_df2D$Status),])


query_df2D <- query_df2D %>%
  mutate(Status = case_when(query_df2D$Sample=="EP3D" & query_df2D$scpred_prediction=="RIS" ~ "RISF"))

ds2df <- nrow(query_df2D[!is.na(query_df2D$Status),])

#Plotting into confusion matrix

table <- matrix(c(ds2dt, ep2df, ds2df, ep2dt),ncol=2,byrow=TRUE)
colnames(table) <- c("DS", "EP")
rownames(table) <- c("DS", "EP")
table <- as.table(table)
table
```


#Can Teo OIS model predict my 3D data?
```{r}
hdfdeb_filt_3D <- subset(hdfbeth_filt, Sample == "DS3D" | Sample == "EP3D")
query3D <- NormalizeData(hdfdeb_filt_3D)
query3D <- scPredict(query3D, Teo_svmradial)

DimHeatmap(query3D, dims=1, cells=500, balanced=TRUE)
DimHeatmap(query3D, dims=1:15, cells=500, balanced=TRUE)
ElbowPlot(query3D)

#Use K-nearest neighbours based on PCA distance in FindNeighbours() function based on previously definited dimensions (10) from elbow plot
query3D <- FindNeighbors(query3D, dims =1:10)

#group cells using findclusters function
query3D <- FindClusters(query3D, resolution = 0.5)
```

```{r}
#Running a UMAP
hdfdeb_filt_3D <- RunUMAP(query3D, dims=1:10)

DimPlot(hdfdeb_filt_3D, reduction = "umap")
tiff("../Results/UMAP_clusters_2D.tiff",width=3300,height=2400,res=400)
DimPlot(hdfdeb_filt_2D, reduction = "umap")


DimPlot(hdfdeb_filt_3D, reduction = "umap",group.by = "Sample")
tiff("../Results/UMAP_samples_2D.tiff",width=3300,height=2400,res=400)
DimPlot(hdfdeb_filt_2D, reduction = "umap",group.by = "Sample")

DimPlot(hdfdeb_filt_3D, reduction = "umap",group.by = "scpred_prediction")
tiff("../Results/UMAP_phase_2D.tiff",width=3300,height=2400,res=400)
DimPlot(hdfdeb_filt_2D, reduction = "umap",group.by = "Phase")

DimPlot(hdfdeb_filt_2D, reduction = "umap", group.by = "scpred_prediction")
```

```{r}
#Combining Teo and Deb's data using seurat integration workflow

# normalise and find variable features seperately

norm_beth <- NormalizeData(hdfbeth_filt)
norm_beth <- FindVariableFeatures(norm_beth, selection.method = "vst", nfeatures = 2000)
norm_beth <- subset(norm_beth, Sample == "DS2D" | Sample == "EP2D" | Sample == "Mix2D")

norm_teo <- NormalizeData(Teo_seurat2)
norm_teo <- FindVariableFeatures(norm_teo, selection.method = "vst", nfeatures = 2000)
norm_teo@meta.data$Sample <- norm_teo@meta.data$Condition2.x
norm_teo@meta.data$Sample[norm_teo@meta.data$Sample == 'Growing'] <- 'EP2D'


# select features that are repeatedly variable across datasets for integration
features <- SelectIntegrationFeatures(object.list = c(norm_beth, norm_teo))

#Find integration anchors
int.nchors <- FindIntegrationAnchors(object.list = c(norm_beth, norm_teo), anchor.features = features)

# this command creates an 'integrated' data assay
beth.teo.combined <- IntegrateData(anchorset = int.nchors)

# specify that we will perform downstream analysis on the corrected data note that the
# original unmodified data still resides in the 'RNA' assay
DefaultAssay(beth.teo.combined) <- "integrated"
```

```{r}
#Selecting highly variable features - high cell to cell variation
beth.teo.combined <- FindVariableFeatures(beth.teo.combined, selection.method ="vst", nfeatures=2000)

#Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(beth.teo.combined), 10)
top10

top2000 <- head(VariableFeatures(beth.teo.combined), 2000)
top2000

#Plot variable features without labels
plot1 <- VariableFeaturePlot(beth.teo.combined)
#Plot variable features with labels
plot2 <- LabelPoints(plot=plot1, points = top10, repel=TRUE)
plot1
plot2
```

#Scaling the data before dimensionality reduction
```{r}
all.genes <- rownames(beth.teo.combined)
beth.teo.combined <- ScaleData(beth.teo.combined, features = all.genes)

#Performing a PCA
beth.teo.combined <- RunPCA(beth.teo.combined, features=VariableFeatures(object = beth.teo.combined))

DimPlot(beth.teo.combined, reduction = "pca")
tiff("../Results/PCA.tiff")
dev.off()


```

## PCA coloured by sample type
```{r}

DimPlot(beth.teo.combined, reduction = "pca", group.by = "Sample") 
tiff("../Results/PCA_grouped.tiff")
DimPlot(beth.teo.combined, reduction = "pca", group.by = "Sample") 

#To determine which PCs to use down the line can use heatmap. Setting 'cells' to a number plots extreme cells on both ends.
```




```{r}
DimHeatmap(beth.teo.combined, dims=1, cells=500, balanced=TRUE)
DimHeatmap(beth.teo.combined, dims=1:15, cells=500, balanced=TRUE)
ElbowPlot(beth.teo.combined)

#Use K-nearest neighbours based on PCA distance in FindNeighbours() function based on previously definited dimensions (10) from elbow plot
beth.teo.combined <- FindNeighbors(beth.teo.combined, dims =1:10)

#group cells using findclusters function
beth.teo.combined <- FindClusters(beth.teo.combined, resolution = 0.5)

#Look at cluster ID's of first 5 cells
head(Idents(beth.teo.combined), 5)
```


# Plot UMAP
```{r}
#Running a UMAP
beth.teo.combined <- RunUMAP(beth.teo.combined, dims=1:10)

DimPlot(beth.teo.combined, reduction = "umap")
tiff("~/Documents/Deb_RNA_seq/Results/UMAP_clusters_All.tiff",width=3300,height=2400,res=400)
DimPlot(beth.teo.combined, reduction = "umap")


DimPlot(beth.teo.combined, reduction = "umap",group.by = "Sample")
tiff("../Results/UMAP_samples.tiff",width=3300,height=2400,res=400)
DimPlot(beth.teo.combined, reduction = "umap",group.by = "Sample")

DimPlot(beth.teo.combined, reduction = "umap",group.by = "Phase")
tiff("../Results/UMAP_phase.tiff",width=3300,height=2400,res=400)
DimPlot(beth.teo.combined, reduction = "umap",group.by = "Phase")